# **R** Point Cloud Processing{#ptcld_analysis}

After running the UAS point cloud [processing script](https://github.com/georgewoolsey/metashape_testing/blob/main/src/software_point_cloud_processing.R) in **R**...the processing tracking data file is used to compare summary statistics on point cloud processing times.

For comparison across software, the SfM point cloud generation processing parameters need are mapped to the Metashape parameters based on the [Pix4D documentation](https://support.pix4d.com/hc/en-us/articles/202557799-Menu-Process-Processing-Options-2-Point-Cloud-and-Mesh-Point-Cloud-PIX4Dmapper) and the [Agisoft Metashape discussion board](https://www.agisoft.com/forum/index.php?topic=11007.0)

```{r processing-dta-ld, results='hide'}
### get tracking data
 # read list of all processed tracking files
  tracking_list_df =
    dplyr::tibble(
      file_full_path = list.files(
          ptcld_processing_dir
          , pattern = ".*_processed_tracking_data\\.csv$"
          , full.names = T, recursive = T
        ) %>% 
        normalizePath()
    ) %>% 
    # get the software used
    dplyr::mutate(
      file_full_path %>% 
        toupper() %>%
        stringr::str_extract_all(pattern = paste(toupper(software_list),collapse = "|"), simplify = T) %>% 
        dplyr::as_tibble() %>% 
        tidyr::unite(col = "software", sep = " ", na.rm = T)
    ) %>% 
    # filter processed tracking files
    dplyr::mutate(
      software = software %>% stringr::word(-1)
      , study_site = file_full_path %>%
        toupper() %>%
        stringr::str_extract(pattern = paste(toupper(study_site_list),collapse = "|"))
      , file_name = file_full_path %>%
        basename() %>%
        stringr::word(1, sep = fixed(".")) %>%
        toupper() %>%
        stringr::str_remove_all("_PROCESSED_TRACKING_DATA")
    ) %>% 
    dplyr::filter(
      !is.na(study_site)
      & study_site %in% toupper(study_site_list)
      & !is.na(software)
      & software %in% toupper(software_list)
    ) %>% 
    # keep only unique files for processing
    dplyr::group_by(software, study_site, file_name) %>%
    dplyr::filter(dplyr::row_number()==1) %>%
    dplyr::ungroup() %>%
    dplyr::rename(tracking_file_full_path = file_full_path)
  
  # tracking_list_df %>% dplyr::glimpse()
  
  # read each tracking data file, bind rows
  ptcld_processing_data = 1:nrow(tracking_list_df) %>%
    purrr::map(function(row_n){
      tracking_list_df %>%
        dplyr::filter(dplyr::row_number() == row_n) %>%
        dplyr::bind_cols(
          read.csv(tracking_list_df$tracking_file_full_path[row_n]) 
        )
    }) %>%
    dplyr::bind_rows()
  
  # ptcld_processing_data %>% dplyr::glimpse()
  # split file name to get processing attributes
    ptcld_processing_data =
      ptcld_processing_data %>%
      tidyr::separate_wider_delim(
        cols = file_name
        , delim = "_"
        , names = paste0(
            "processing_attribute"
            , 1:(max(stringr::str_count(ptcld_processing_data$file_name, "_"))+1)
          )
        , too_few = "align_start"
        , cols_remove = F
      ) %>%
      # not sure how to map processing attributes for pix4d and opendronemap ??????????????
      dplyr::mutate(
        # temporary
        qqq = dplyr::case_when(
            tolower(software) == "pix4d" ~ processing_attribute2
            , T ~ processing_attribute1
          )
        , fff = dplyr::case_when(
            tolower(software) == "pix4d" ~ processing_attribute3
            , T ~ processing_attribute2
          )
        # mapping 
        , depth_maps_generation_quality = dplyr::case_when(
            tolower(qqq) %in% c("ultrahigh", "ultra", "original", "origianl") ~ "ultra high"
            , tolower(qqq) %in% c("half") ~ "high"
            , tolower(qqq) %in% c("quarter") ~ "medium"
            , tolower(qqq) %in% c("eighth","eightht") ~ "low"
            , T ~ tolower(qqq)
          ) %>% 
          factor(
            ordered = TRUE
            , levels = c(
              "lowest"
              , "low"
              , "medium"
              , "high"
              , "ultra high"
            )
          ) %>% forcats::fct_rev()
        , depth_maps_generation_filtering_mode = dplyr::case_when(
            tolower(fff) %in% c("high") & 
              tolower(software) %in% c("opendronemap") ~ "disabled"
            , tolower(fff) %in% c("high")
              & tolower(software) %in% c("pix4d") ~ "disabled"
            , tolower(fff) %in% c("medium")
              & tolower(software) %in% c("opendronemap") ~ "mild"
            , tolower(fff) %in% c("optimal")
              & tolower(software) %in% c("pix4d") ~ "mild"
            , tolower(fff) %in% c("low")
              & tolower(software) %in% c("opendronemap") ~ "moderate"
            , tolower(fff) %in% c("low")
              & tolower(software) %in% c("pix4d") ~ "moderate"
            , tolower(fff) %in% c("lowest")
              & tolower(software) %in% c("opendronemap") ~ "aggressive"
            , T ~ tolower(fff)
          ) %>% 
          factor(
            ordered = TRUE
            , levels = c(
              "disabled"
              , "mild"
              , "moderate"
              , "aggressive"
            )
          ) %>% forcats::fct_rev()
      )
```

what have we done?

```{r}
ptcld_processing_data %>% dplyr::glimpse()
```

```{r, include=F, eval=F}
ptcld_processing_data %>% 
  dplyr::filter(software == "OPENDRONEMAP" & 
    depth_maps_generation_filtering_mode %in% c("aggressive", "disabled")
                  ) %>% 
  ggplot(aes(x = depth_maps_generation_quality, y = number_of_points, color = depth_maps_generation_filtering_mode)) + 
  geom_point() +
  scale_y_log10() +
  facet_wrap(facets = vars(study_site), ncol = 2)
```


what is this mapping?

```{r}
# quality
ptcld_processing_data %>% 
  dplyr::count(depth_maps_generation_quality, qqq, software) %>% 
  ggplot(aes(x = tolower(software), y = depth_maps_generation_quality, label = tolower(qqq))) +
    geom_tile(fill = NA, color = "black") +
    ggrepel::geom_text_repel(color = "gray33") +
    labs(y = "Mapped: Depth Map Quality", x = "") +
    scale_x_discrete(position = "top") +
    coord_cartesian(expand = F) +
    theme_light() +
    theme(
      panel.grid = element_blank()
      , axis.text = element_text(size = 11, face = "bold", color = "black")
      , panel.border = element_rect(color = "black")
    )
ggplot2::ggsave("../data/mapped_quality.jpg", height = 8, width = 6)
```

```{r}
# filtering
ptcld_processing_data %>% 
  dplyr::count(depth_maps_generation_filtering_mode, fff, software) %>% 
  ggplot(aes(x = tolower(software), y = depth_maps_generation_filtering_mode, label = tolower(fff))) +
    geom_tile(fill = NA, color = "black") +
    ggrepel::geom_text_repel(color = "gray33") +
    labs(y = "Mapped: Filtering Mode", x = "") +
    scale_x_discrete(position = "top") +
    coord_cartesian(expand = F) +
    theme_light() +
    theme(
      panel.grid = element_blank()
      , axis.text = element_text(size = 11, face = "bold", color = "black")
      , panel.border = element_rect(color = "black")
    )
ggplot2::ggsave("../data/mapped_filtering_mode.jpg", height = 8, width = 6)
```

```{r}
# !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! Filtering
# !!!! keep only one kind of pix4d, all metashape and odm ???????
ptcld_processing_data = ptcld_processing_data %>% 
  dplyr::select(-c(qqq,fff)) %>% 
  dplyr::filter(
    dplyr::case_when(
      tolower(software) == "pix4d" & tolower(processing_attribute1) == "original" ~ T
      , tolower(software) != "pix4d" ~ T
      , T ~ F
    ) == T
  )
```

## Number of files summary

```{r, fig.height=8}
ptcld_processing_data %>% 
  dplyr::count(software, depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
  ggplot(mapping = aes(
    y = depth_maps_generation_quality
    , x = depth_maps_generation_filtering_mode
    , fill = n
    , label = n
  )) +
  geom_tile(color = "white") +
  geom_text(color = "white", size = 3) +
  facet_grid(cols = vars(software)) + 
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_viridis_c(option = "mako", direction=-1, begin = 0.2, end = 0.8) +
  labs(
    x = "filtering mode"
    , y = "depth map quality"
    , fill = "number of sites"
  ) +
  theme_light() + 
  theme(
    legend.position = "none"
    , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
    , panel.background = element_blank()
    , panel.grid = element_blank()
    , plot.subtitle = element_text(hjust = 0.5)
    , strip.text = element_text(color = "black", face = "bold")
  )

ggplot2::ggsave("../data/n_sites_comp_quick.png", height = 9, width = 8)
```

## Processing Time Summary

Total processing time by depth map generation quality and depth map filtering mode 

```{r processing-dta-tot}
ptcld_processing_data %>% 
  ggplot(
    mapping = aes(
      x = depth_maps_generation_quality
      , y = timer_total_time_mins
      , color = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
  geom_boxplot(alpha = 0.6) +
  scale_color_viridis_d(option = "plasma") +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_log10(
    labels = scales::comma_format(suffix = " mins", accuracy = 1)
    , breaks = scales::breaks_log(n = 9)
  ) +
  labs(
    color = "Filtering Mode"
    , fill = "Filtering Mode"
    , y = "Point Cloud Total Processing Time"
    , x = "Quality"
    , title = bquote(
        bold("R") ~
        "point cloud total processing time by depth map generation quality and filtering mode"
    )
    , caption = "*Note the log scale on the y-axis"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 15, size = 6, alpha = 0.9))
  )
```

Notice there are some outlier study sites in the point cloud processing time

```{r processing-dta-tot-area}
ptcld_processing_data %>% 
  ggplot(
    mapping = aes(
      y = timer_total_time_mins
      , x = depth_maps_generation_quality
      , color = depth_maps_generation_filtering_mode
    )
  ) +
  geom_point(size = 3, alpha = 0.8) +
  facet_grid(
    cols = vars(study_site)
    , labeller = label_wrap_gen(width = 35, multi_line = TRUE)
  ) +
  scale_color_viridis_d(option = "plasma") +
  scale_y_log10(
    labels = scales::comma_format(suffix = " mins", accuracy = 1)
    , breaks = scales::breaks_log(n = 9)
  ) +
  labs(
    color = "Filtering Mode"
    , y = "Point Cloud Total Processing Time"
    , x = "Quality"
    , title = bquote(
        bold("R") ~
        "point cloud total processing time by depth map generation quality and filtering mode"
    )
    , subtitle = "by Study Site"
    , caption = "*Note the log scale on the y-axis"
  ) +
  theme_light() + 
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , strip.text = element_text(color = "black", face = "bold")
    , axis.text.x = element_text(angle = 90)
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 15, size = 6, alpha = 0.9))
  )
```

## Processing Time vs # Points

```{r processing-dta-time-pts}
ptcld_processing_data %>% 
  ggplot(
    mapping = aes(
      x = number_of_points
      , y = timer_total_time_mins
    )
  ) +
  geom_point(alpha = 0.7, color = "navy") +
  scale_y_log10(
    labels = scales::comma_format(suffix = " mins", accuracy = 1)
    , breaks = scales::breaks_log(n = 9)
  ) +
  scale_x_log10(
    labels = scales::comma_format(suffix = " M", scale = 1e-6, accuracy = 1)
    , breaks = scales::breaks_log(n = 6)
  ) +
  labs(
    y = "Point Cloud Total Processing Time"
    , x = "Dense Point Cloud # Points"
    , title = bquote(
        bold("R") ~
        "point cloud total processing time versus dense point cloud number of points"
    )
    , caption = "*Note the log scale on both axes"
  ) +
  theme_light()
```

## Processing Section Timing

```{r time-breakdown}
ptcld_processing_data %>% 
  dplyr::select(
    depth_maps_generation_quality
    , tidyselect::ends_with("_mins")
  ) %>% 
  dplyr::select(-c(timer_total_time_mins)) %>% 
  tidyr::pivot_longer(
    cols = -c(depth_maps_generation_quality)
    , names_to = "section"
    , values_to = "mins"
  ) %>% 
  # dplyr::count(depth_maps_generation_quality, section)
  dplyr::group_by(depth_maps_generation_quality, section) %>% 
  dplyr::summarise(med_mins = median(mins)) %>% 
  dplyr::group_by(depth_maps_generation_quality) %>% 
  dplyr::mutate(
    total_mins = sum(med_mins)
    , pct_mins = med_mins/total_mins 
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    section = section %>% 
      stringr::str_remove_all("timer_") %>% 
      stringr::str_remove_all("_time_mins") %>% 
      factor(
        ordered = T
        , levels = c(
          "tile"
          , "class_dtm_norm_chm"
          , "treels"
          , "itd"
          , "estdbh"
          , "competition"
          , "silv"
          ## olde
          # "tile"
          # , "denoise"
          # , "classify"
          # , "dtm"
          # , "normalize"
          # , "chm"
          # , "treels"
          # , "itd"
          # , "estdbh"
          # , "competition"
          # , "silv"
        )
        , labels = c(
          "Tile"
          , "Classify+Denoise+DTM+Normalize+CHM"
          , "TreeLS SfM DBH"
          , "CHM I.T.D."
          , "Local DBH Est."
          , "Tree Competition"
          , "Silvicultural Metrics"
        )
      ) %>% forcats::fct_rev()
  ) %>%
  
ggplot(
  mapping = aes(x = pct_mins, y = depth_maps_generation_quality, fill=section, group=section)
) +
  geom_col(
    width = 0.7, alpha=0.8
  ) +
  geom_text(
    mapping = aes(
        label = scales::percent(ifelse(pct_mins>=0.06,pct_mins,NA), accuracy = 1)
        , fontface = "bold"
      )
    , position = position_stack(vjust = 0.5)
    , color = "black", size = 4
  ) +
  scale_fill_viridis_d(option = "turbo", begin = 0.1, end = 0.9) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(
    fill = "R script\nsection"
    , y = "depth map quality"
    , x = "% Point Cloud Total Processing Time"
    , title = bquote(
        bold("R") ~
        "point cloud total processing time by depth map generation quality and R script section"
    )
    , subtitle = "Median across software, study site, & depth map filtering mode "
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , legend.title = element_text(size=7)
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.title.y = element_text(size = 8)
    , axis.text.x = element_blank()
    , axis.text.y = element_text(color = "black",size=10, face = "bold")
    , axis.ticks.x = element_blank()
  ) +
  guides(
    fill = guide_legend(nrow = 3, byrow = T, reverse = T, override.aes = list(alpha = 0.9))
  )  

ggplot2::ggsave("../data/processing_time_brkdown.png", width = 8.5, height = 6)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

