# Statistical Analysis: Validation{#stats_validation}

In this section, we'll evaluate the influence of the processing parameters on UAS-derived tree detection and monitoring. 

The UAS and Field validation data was built and described in [this section](#field_valid).

The objective of this study is to determine the influence of different structure from motion (SfM) software (e.g. Agisoft  Metashap, OpenDroneMap, Pix4D) and processing parameters on F-score which is a measure of overall tree detection performance.

All of the predictor variables of interest in this study are categorical (a.k.a. factor or nominal) while the predicted variables are metric and include F-score (ranges from 0-1) and error (e.g. MAPE, RMSE). This type of statistical analysis is described in the second edition of Kruschke's [*Doing Bayesian data analysis* (2015)](https://sites.google.com/site/doingbayesiandataanalysis/) and here we will build a Bayesian approach based on [Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/). This analysis was greatly enhanced by [A. Solomon Kurz's ebook supplement](https://solomonkurz.netlify.app/book/) to [Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/).

For a more in-depth review of the traditional treatment of this sort of data structure called multifactor analysis of variance (ANOVA) compared to the Bayesian hierarchical generalization of the traditional ANOVA model used here see [this previous section](#stats_processing_time).

## Summary Statistics

What is this data?

```{r}
# load data if needed
if(ls()[ls() %in% "ptcld_validation_data"] %>% length()==0){
 ptcld_validation_data = readr::read_csv("../data/ptcld_full_analysis_data.csv") %>% 
   dplyr::mutate(
      depth_maps_generation_quality = factor(
          depth_maps_generation_quality %>% 
            tolower() %>% 
            stringr::str_replace_all("ultrahigh", "ultra high")
          , ordered = TRUE
          , levels = c(
            "lowest"
            , "low"
            , "medium"
            , "high"
            , "ultra high"
          )
        ) %>% forcats::fct_rev()
      , depth_maps_generation_filtering_mode = factor(
          depth_maps_generation_filtering_mode %>% tolower()
          , ordered = TRUE
          , levels = c(
            "disabled"
            , "mild"
            , "moderate"
            , "aggressive"
          )
        ) %>% forcats::fct_rev()
    )
}
# what is this data?
ptcld_validation_data %>% dplyr::glimpse()
# a row is unique by...
identical(
  nrow(ptcld_validation_data)
  , ptcld_validation_data %>% 
    dplyr::distinct(
      study_site, software
      , depth_maps_generation_quality
      , depth_maps_generation_filtering_mode
      , processing_attribute3 # need to align all by software so this will go away or be filled
    ) %>% 
    nrow()
)

```

Summary by metrics of interest

```{r}
sum_stats_dta = function(my_var){
  sum_fns <- list(
    n = ~sum(ifelse(is.na(.x), 0, 1))  
    , min = ~min(.x, na.rm = TRUE)
    , max = ~max(.x, na.rm = TRUE)
    , mean = ~mean(.x, na.rm = TRUE)
    , median = ~median(.x, na.rm = TRUE)
    , sd = ~sd(.x, na.rm = TRUE)
  )
  # plot
  print(
    ggplot(
      data = ptcld_validation_data %>% 
        dplyr::group_by(.data[[my_var]]) %>%
        dplyr::mutate(m = median(f_score))
      , mapping = aes(
        y = .data[[my_var]]
        , x = f_score, fill = m)
      ) +
      geom_violin(color = NA) + 
      geom_boxplot(width = 0.1, outlier.shape = NA, fill = NA, color = "black") +
      geom_rug() +
      scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.9, direction = -1) +
      labs(
        x = "F-score"
        , y = stringr::str_replace_all(my_var, pattern = "_", replacement = " ")
        , subtitle = stringr::str_replace_all(my_var, pattern = "_", replacement = " ") %>% 
          stringr::str_to_title()
      ) +
      theme_light() +
      theme(legend.position = "none")
  )
  # summarize data
  (
  ptcld_validation_data %>% 
    dplyr::group_by(dplyr::across(dplyr::all_of(my_var))) %>%
    dplyr::summarise(
      dplyr::across(f_score, sum_fns)
      , .groups = 'drop_last'
    ) %>% 
    kableExtra::kbl() %>% 
    kableExtra::kable_styling()
  )
}
# sum_stats_dta("software")
```

summarize for all variables of interest

```{r}
c("software", "study_site"
  , "depth_maps_generation_quality"
  , "depth_maps_generation_filtering_mode"
) %>% 
  purrr::map(sum_stats_dta)
```

## One Nominal Predictor{#f_one_pred_mod}

We'll start by exploring the influence of the depth map generation quality parameter on the SfM-derived tree detection performance based on the F-score.

### Summary Statistics

Summary statistics by group:

```{r}
ptcld_validation_data %>% 
  dplyr::group_by(depth_maps_generation_quality) %>% 
  dplyr::summarise(
    mean_f_score = mean(f_score, na.rm = T)
    # , med_f_score = median(f_score, na.rm = T)
    , sd_f_score = sd(f_score, na.rm = T)
    , n = dplyr::n()
  ) %>% 
  kableExtra::kbl(digits = 2, caption = "summary statistics: F-score by depth map quality") %>% 
  kableExtra::kable_styling()
```

### Bayesian{#f_one_pred_mod_bays}

[Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/) notes: 

> The terminology, "analysis of variance," comes from a decomposition of overall data variance into within-group variance and between-group variance (Fisher, 1925). Algebraically, the sum of squared deviations of the scores from their overall mean equals the sum of squared deviations of the scores from their respective group means plus the sum of squared deviations of the group means from the overall mean. In other words, the total variance can be partitioned into within-group variance plus between-group variance. Because one definition of the word "analysis" is separation into constituent parts, the term ANOVA accurately describes the underlying algebra in the traditional methods. That algebraic relation is not used in the hierarchical Bayesian approach presented here. The Bayesian method can estimate component variances, however. Therefore, the Bayesian approach is not ANOVA, but is analogous to ANOVA. (p. 556)

and see section 19 from [Kurz's ebook supplement](https://bookdown.org/content/3686/metric-predicted-variable-with-one-nominal-predictor.html)

The metric predicted variable with one nominal predictor variable model has the form:

\begin{align*}
y_{i}  &\sim {\sf Normal} \bigl(\mu_{i}, \sigma_{y} \bigr) \\
\mu_{i} &= \beta_0 + \sum_{j=1}^{J} \beta_{1[j]} x_{1[j]} \bigl(i\bigr) \\
\beta_{0}  &\sim {\sf Normal} (0,10) \\ 
\beta_{1[j]}  &\sim {\sf Normal} (0,\sigma_{\beta_{1}}) \\ 
\sigma_{\beta_{1}} &\sim {\sf uniform} (0,100) \\ 
\sigma_{y} &\sim {\sf uniform} (0,100) \\ 
\end{align*}

, where $j$ is the depth map generation quality setting corresponding to observation $i$

*to start, we'll use the default `brms::brm` prior settings which may not match those described in the model specification above* 

```{r}
brms_f_mod1 = brms::brm(
  formula = f_score ~ 1 + (1 | depth_maps_generation_quality)
  , data = ptcld_validation_data
  , family = brms::brmsfamily(family = "gaussian")
  , iter = 4000, warmup = 2000, chains = 4
  , cores = round(parallel::detectCores()/2)
  , file = paste0(rootdir, "/fits/brms_f_mod1")
)
```

check the trace plots for problems with convergence of the Markov chains

```{r}
plot(brms_f_mod1)
```

check the prior distributions

```{r}
# check priors
brms::prior_summary(brms_f_mod1) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling()
```

The `brms::brm` model summary

```{r}
brms_f_mod1 %>% 
  brms::posterior_summary() %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>%
  dplyr::rename_with(tolower) %>% 
  dplyr::filter(
    stringr::str_starts(parameter, "b_") 
    | stringr::str_starts(parameter, "r_") 
    | parameter == "sigma"
  ) %>%
  dplyr::mutate(
    parameter = parameter %>% 
      stringr::str_remove_all("b_depth_maps_generation_quality") %>% 
      stringr::str_remove_all("r_depth_maps_generation_quality")
  ) %>% 
  kableExtra::kbl(digits = 2, caption = "Bayesian one nominal predictor: F-score by depth map quality") %>% 
  kableExtra::kable_styling()
```

With the `stats::coef` function, we can get the group-level summaries in a "non-deflection" metric. In the model, the group means represented by $\beta_{1[j]}$ are deflections from overall baseline, such that the deflections sum to zero (see [Kruschke (2015, p.554)](https://sites.google.com/site/doingbayesiandataanalysis/)). Summaries of the group-specific deflections are available via the `brms::ranef` function.

```{r}
stats::coef(brms_f_mod1) %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "group") %>% 
  dplyr::rename_with(
    .cols = -c("group")
    , .fn = ~ stringr::str_remove_all(.x, "depth_maps_generation_quality.")
  ) %>% 
  kableExtra::kbl(digits = 2, caption = "brms::brm model: F-score by depth map quality") %>% 
  kableExtra::kable_styling()
```

We can look at the model noise standard deviation $\sigma_y$

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod1) %>% 
# plot
  ggplot(aes(x = sigma, y = 0)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21, point_size = 3
    , quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(latex2exp::TeX("$\\sigma_y$")) +
  theme_light()
```

plot the posterior distributions of the conditional means with the median F-score and the 95% highest posterior density interval (HDI)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(brms_f_mod1) %>% 
  dplyr::mutate(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_quality
      , fill = depth_maps_generation_quality
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "inferno", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "depth map quality", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

we can also make pairwise comparisons

```{r}
# first we need to define the contrasts to make
contrast_list = 
  tidyr::crossing(
    x1 = unique(ptcld_validation_data$depth_maps_generation_quality)
    , x2 = unique(ptcld_validation_data$depth_maps_generation_quality)
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      dplyr::everything()
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
  ) %>% 
  dplyr::filter(x1<x2) %>% 
  dplyr::arrange(x1,x2) %>% 
  dplyr::mutate(dplyr::across(dplyr::everything(), as.character)) %>% 
  purrr::transpose()

# contrast_list

# obtain posterior draws and calculate contrasts using tidybayes::compare_levels
brms_contrast_temp = 
  brms_f_mod1 %>% 
    tidybayes::spread_draws(r_depth_maps_generation_quality[depth_maps_generation_quality]) %>% 
    dplyr::mutate(
      depth_maps_generation_quality = depth_maps_generation_quality %>% 
        stringr::str_replace_all("\\.", " ") %>% 
        factor(
          levels = levels(ptcld_validation_data$depth_maps_generation_quality)
          , ordered = T
        )
    ) %>% 
    dplyr::rename(value = r_depth_maps_generation_quality) %>% 
    tidybayes::compare_levels(
      value
      , by = depth_maps_generation_quality
      , comparison = contrast_list
        # tidybayes::emmeans_comparison("revpairwise") 
        #"pairwise"
    ) 

# generate the contrast column for creating an ordered factor
brms_contrast_temp =
  brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = depth_maps_generation_quality
    , delim = " - "
    , names = paste0(
        "sorter"
        , 1:(max(stringr::str_count(brms_contrast_temp$depth_maps_generation_quality, "-"))+1)
      )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = depth_maps_generation_quality %>% forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
        as.numeric()
      )
  )

# median_hdi summary for coloring 
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::group_by(contrast) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
    factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )
# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot, finally
brms_contrast_temp %>% 
  ggplot(aes(x = value, y = contrast, fill = sig_level)) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = c(0.5,0.95)
      # , slab_fill = "gray22", slab_alpha = 1
      , interval_color = "black", point_color = "black", point_fill = "black"
      , justification = -0.01
    ) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
    scale_fill_viridis_d(
      option = "mako", begin = 0.3
      , drop = F
      # , labels = scales::percent
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "depth map quality"
      , x = "constrast (F-score)"
      , fill = "Pr(contrast > 0)"
      , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts"
    ) +
    theme_light() +
    theme(legend.position = "top", legend.direction = "horizontal") +
    guides(
      fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    )
```

and summarize these contrasts

```{r}
# # can also use the following as substitute for the "tidybayes::spread_draws" used above to get same result
brms_contrast_temp %>% 
  dplyr::group_by(contrast) %>% 
  tidybayes::median_hdi(value) %>% 
  select(-c(.point,.interval, .width)) %>% 
  dplyr::arrange(desc(contrast)) %>% 
  dplyr::rename(difference=value) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts in F-score"
    , col.names = c(
      "quality contrast"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Two Nominal Predictors{#f_two_pred_mod}

Now, we'll determine the combined influence of the depth map generation quality and the depth map filtering parameters on the SfM-derived tree detection performance based on the F-score.

### Summary Statistics

Summary statistics by group:

```{r}
ptcld_validation_data %>% 
  dplyr::group_by(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
  dplyr::summarise(
    mean_f_score = mean(f_score, na.rm = T)
    # , med_f_score = median(f_score, na.rm = T)
    , sd_f_score = sd(f_score, na.rm = T)
    , n = dplyr::n()
  ) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "summary statistics: F-score by depth map quality and filtering mode"
    , col.names = c(
      "depth map quality"
      , "filtering mode"
      , "mean f-score"
      , "sd"
      , "n"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```

### Bayesian{#f_two_pred_mod_bays}

[Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/) describes the Hierarchical Bayesian approach to describe groups of metric data with multiple nominal predictors: 

> This chapter considers data structures that consist of a metric predicted variable and two (or more) nominal predictors....The traditional treatment of this sort of data structure is called multifactor analysis of variance (ANOVA). Our Bayesian approach will be a hierarchical generalization of the traditional ANOVA model. The chapter also considers generalizations of the traditional models, because it is straight forward in Bayesian software to implement heavy-tailed distributions to accommodate outliers, along with hierarchical structure to accommodate heterogeneous variances in the different groups. (pp. 583–584)

and see section 20 from [Kurz's ebook supplement](https://bookdown.org/content/3686/metric-predicted-variable-with-multiple-nominal-predictors.html)

The metric predicted variable with two nominal predictor variables model has the form:

\begin{align*}
y_{i}  &\sim {\sf Normal} \bigl(\mu_{i}, \sigma_{y} \bigr) \\
\mu_{i} &= \beta_0 + \sum_{j} \beta_{1[j]} x_{1[j]} + \sum_{k} \beta_{2[k]} x_{2[k]}  + \sum_{j,k} \beta_{1\times2[j,k]} x_{1\times2[j,k]} \\
\beta_{0}  &\sim {\sf Normal} (0,100) \\ 
\beta_{1[j]}  &\sim {\sf Normal} (0,\sigma_{\beta_{1}}) \\ 
\beta_{2[k]}  &\sim {\sf Normal} (0,\sigma_{\beta_{2}}) \\ 
\beta_{1\times2[j,k]}  &\sim {\sf Normal} (0,\sigma_{\beta_{1\times2}}) \\ 
\sigma_{\beta_{1}} &\sim {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{2}} &\sim {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{1\times2}} &\sim {\sf Gamma} (1.28,0.005) \\ 
\sigma_{y} &\sim {\sf Cauchy} (0,109) \\ 
\end{align*}

, where $j$ is the depth map generation quality setting corresponding to observation $i$ and $k$ is the depth map filtering mode setting corresponding to observation $i$

*for this model, we'll define the priors following [Kurz](https://bookdown.org/content/3686/metric-predicted-variable-with-multiple-nominal-predictors.html#implementation-in-jags-brms.-1)* who notes that:

>The noise standard deviation $\sigma_y$ is depicted in the prior statement including the argument `class = sigma`...in order to be weakly informative, we will use the half-Cauchy. Recall that since the brms default is to set the lower bound for any variance parameter to 0, there’s no need to worry about doing so ourselves. So even though the syntax only indicates `cauchy`, it’s understood to mean Cauchy with a lower bound at zero; since the mean is usually 0, that makes this a half-Cauchy...The tails of the half-Cauchy are sufficiently fat that, in practice, I’ve found it doesn’t matter much what you set the $SD$ of its prior to.

```{r}
# from Kurz: 
gamma_a_b_from_omega_sigma <- function(mode, sd) {
  if (mode <= 0) stop("mode must be > 0")
  if (sd   <= 0) stop("sd must be > 0")
  rate <- (mode + sqrt(mode^2 + 4 * sd^2)) / (2 * sd^2)
  shape <- 1 + mode * rate
  return(list(shape = shape, rate = rate))
}

mean_y_temp <- mean(ptcld_validation_data$f_score)
sd_y_temp   <- sd(ptcld_validation_data$f_score)

omega_temp <- sd_y_temp / 2
sigma_temp <- 2 * sd_y_temp

s_r_temp <- gamma_a_b_from_omega_sigma(mode = omega_temp, sd = sigma_temp)

stanvars_temp <- 
  brms::stanvar(mean_y_temp,    name = "mean_y") + 
  brms::stanvar(sd_y_temp,      name = "sd_y") +
  brms::stanvar(s_r_temp$shape, name = "alpha") +
  brms::stanvar(s_r_temp$rate,  name = "beta")
```

Now fit the model.

```{r}
brms_f_mod2 = brms::brm(
  formula = f_score ~ 1 + 
    (1 | depth_maps_generation_quality) +
    (1 | depth_maps_generation_filtering_mode) + 
    (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode)
  , data = ptcld_validation_data
  , family = brms::brmsfamily(family = "gaussian")
  , iter = 6000, warmup = 3000, chains = 4
  , cores = round(parallel::detectCores()/2)
  , prior = c(
    brms::prior(normal(mean_y, sd_y * 5), class = "Intercept")
    , brms::prior(gamma(alpha, beta), class = "sd")
    , brms::prior(cauchy(0, sd_y), class = "sigma")
  )
  , stanvars = stanvars_temp
  , file = paste0(rootdir, "/fits/brms_f_mod2")
)
```

check the trace plots for problems with convergence of the Markov chains

```{r, fig.height=8}
plot(brms_f_mod2)
```

check the prior distributions

```{r}
# check priors
brms::prior_summary(brms_f_mod2) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling()
```

The `brms::brm` model summary

```{r}
brms_f_mod2 %>% 
  brms::posterior_summary() %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>%
  dplyr::rename_with(tolower) %>% 
  dplyr::filter(
    stringr::str_starts(parameter, "b_") 
    | stringr::str_starts(parameter, "r_") 
    | stringr::str_starts(parameter, "sd_") 
    | parameter == "sigma"
  ) %>%
  dplyr::mutate(
    parameter = parameter %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering")
  ) %>% 
  kableExtra::kbl(digits = 2, caption = "Bayesian two nominal predictors: F-score by depth map quality and filtering mode") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```


We can look at the model noise standard deviation $\sigma_y$

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod2) %>% 
# plot
  ggplot(aes(x = sigma, y = 0)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21, point_size = 3
    , quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(latex2exp::TeX("$\\sigma_y$")) +
  theme_light()
```

how is it compared to the first model?

```{r}
# how is it compared to the first model
dplyr::bind_rows(
    brms::as_draws_df(brms_f_mod1) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "one nominal predictor")
    , brms::as_draws_df(brms_f_mod2) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "two nominal predictor")
  ) %>% 
  dplyr::relocate(model) %>% 
  kableExtra::kbl(digits = 2, caption = "brms::brm model noise standard deviation comparison") %>% 
    kableExtra::kable_styling() %>% 
    kableExtra::scroll_box(width = "8in", height = "6in")
```

plot the posterior distributions of the conditional means with the median F-score and the 95% highest posterior density interval (HDI)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(brms_f_mod2, allow_new_levels = T) %>% 
  dplyr::rename(value = .epred) %>% 
  dplyr::mutate(depth_maps_generation_quality = depth_maps_generation_quality %>% forcats::fct_rev()) %>% 
  # plot
  ggplot(
    mapping = aes(
      y = value, x = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_eye(
      point_interval = median_hdi, .width = .95
      , slab_alpha = 0.8
      , interval_color = "black", linewidth = 1
      , point_color = "black", point_fill = "black", point_size = 1
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_y_continuous(breaks = scales::extended_breaks(n=10)) +
    facet_grid(cols = vars(depth_maps_generation_quality)) +
    labs(
      x = "filtering mode", y = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
      , fill = "Filtering Mode"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , legend.direction  = "horizontal"
      , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
      , strip.text = element_text(color = "black", face = "bold")
    ) 
    # guides(
    #   fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    # )
```

we can also make pairwise comparisons

```{r}
brms_contrast_temp =
  ptcld_validation_data %>%
    dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
    tidybayes::add_epred_draws(brms_f_mod2, allow_new_levels = T) %>% 
    dplyr::rename(value = .epred) %>% 
    tidybayes::compare_levels(
      value
      , by = depth_maps_generation_quality
      , comparison = 
        contrast_list
        # tidybayes::emmeans_comparison("revpairwise") 
        #"pairwise"
    ) %>% 
    dplyr::rename(contrast = depth_maps_generation_quality)

# separate contrast
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = contrast
    , delim = " - "
    , names = paste0(
        "sorter"
        , 1:(max(stringr::str_count(brms_contrast_temp$contrast, "-"))+1)
      )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = contrast %>% 
      forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
        as.numeric()
      )
    , depth_maps_generation_filtering_mode = depth_maps_generation_filtering_mode %>% 
      factor(
        levels = levels(ptcld_validation_data$depth_maps_generation_filtering_mode)
        , ordered = T
      )
  ) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast, depth_maps_generation_filtering_mode) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
    factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      , fill = sig_level # pct_gt_zero
      # , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = c(0.5,0.95)
      # , slab_fill = "gray22", slab_alpha = 1
      , interval_color = "black", point_color = "black", point_fill = "black"
      , justification = -0.01
    ) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
    scale_fill_viridis_d(
      option = "mako", begin = 0.3
      , drop = F
      # , labels = scales::percent
    ) +
    # scale_fill_viridis_c(
    #   option = "mako", begin = 0.3, direction = -1
    #   , labels = scales::percent
    # ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    facet_grid(cols = vars(depth_maps_generation_filtering_mode)) +
    labs(
      y = "depth map quality"
      , x = "constrast (f-score)"
      , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby filtering mode"
      , fill = "Pr(contrast > 0)"
    ) +
    theme_light() +
    theme(
      legend.position = "top"
      , legend.direction  = "horizontal"
      , strip.text = element_text(color = "black", face = "bold")
    ) +
    guides(
      fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    )
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, depth_maps_generation_filtering_mode) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, depth_maps_generation_filtering_mode) %>% 
  dplyr::select(contrast, depth_maps_generation_filtering_mode, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "quality contrast"
      , "filtering mode"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```

[Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/) notes that for the multiple nominal predictors model: 

> In applications with multiple levels of the factors, it is virtually always the case that we are interested in comparing particular levels with each other...These sorts of comparisons, which involve levels of a single factor and collapse across the other factor(s), are called main effect comparisons or contrasts.(p. 595)

First, let's collapse across the filtering mode to compare the depth map quality setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod2
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_quality
      , fill = depth_maps_generation_quality
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "inferno", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "depth map quality", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [one nominal predictor model above](#f_one_pred_mod_bays)

```{r}
# let's compare these results to the results from our [one nominal predictor model above](#one_pred_mod)
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod2
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod2") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(brms_f_mod1) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod1")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(depth_maps_generation_quality), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[1:2]) +
  labs(
    y = "", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

these results are as expected, with [Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/) noting:

> It is important to realize that the estimates of interaction contrasts are typically much more uncertain than the estimates of simple effects or main effects...This large uncertainty of an interaction contrast is caused by the fact that it involves at least four sources of uncertainty (i.e., at least four groups of data), unlike its component simple effects which each involve only half of those sources of uncertainty. In general, interaction contrasts require a lot of data to estimate accurately. (p. 598)

For completeness, let's also collapse across the depth map quality to compare the filtering mode setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod2
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "filtering mode", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

...it looks like the variation in F-score is driven by the depth map quality setting. We can quantify the variation in F-score by comparing the $\sigma$ posteriors.

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod2) %>% 
  dplyr::select(c(sigma,tidyselect::starts_with("sd_"))) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  # dplyr::group_by(name) %>% 
  # tidybayes::median_hdi(value) %>% 
  dplyr::mutate(
    name = name %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering") %>% 
      forcats::fct_reorder(value)
  ) %>%
# plot
  ggplot(aes(x = value, y = name)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21 #, point_size = 3
    , quantiles = 100
  ) +
  labs(x = "", y = "") +
  theme_light()
```

Finally we can perform model selection via information criteria, from section 10 in [Kurz's ebook supplement](https://bookdown.org/content/3686/metric-predicted-variable-with-one-nominal-predictor.html):

> expected log predictive density (`elpd_loo`), the estimated effective number of parameters (`p_loo`), and the Pareto smoothed importance-sampling leave-one-out cross-validation (PSIS-LOO; `looic`). Each estimate comes with a standard error (i.e., `SE`). Like other information criteria, the LOO values aren't of interest in and of themselves. However, the estimate of one model's LOO relative to that of another can be of great interest. We generally prefer models with lower information criteria. With the `brms::loo_compare()` function, we can compute a formal difference score between two models...The `brms::loo_compare()` output rank orders the models such that the best fitting model appears on top.

```{r}
brms_f_mod1 = brms::add_criterion(brms_f_mod1, criterion = c("loo", "waic"))
brms_f_mod2 = brms::add_criterion(brms_f_mod2, criterion = c("loo", "waic"))
brms::loo_compare(brms_f_mod1, brms_f_mod2, criterion = "loo")
# brms::model_weights(brms_f_mod1, brms_f_mod2) %>% round()
```

These models are not significantly different suggesting that the filtering mode is not contributing much to the variation in SfM-derived tree detection reliability after accounting for the quality setting

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Three Nominal Predictors{#f_three_pred_mod}

Now, we'll add the SfM processing software to our model which includes the depth map generation quality and the depth map filtering parameters to quantify the SfM-derived tree detection performance based on the F-score.

### Summary Statistics

Summary statistics by group:

```{r}
ptcld_validation_data %>% 
  dplyr::group_by(depth_maps_generation_quality, depth_maps_generation_filtering_mode, software) %>%
  dplyr::summarise(
    mean_f_score = mean(f_score, na.rm = T)
    # , med_f_score = median(f_score, na.rm = T)
    , sd_f_score = sd(f_score, na.rm = T)
    , n = dplyr::n()
  ) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "summary statistics: F-score by depth map quality, filtering mode, and software"
    , col.names = c(
      "depth map quality"
      , "filtering mode"
      , "software"
      , "mean f-score"
      , "sd"
      , "n"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "8in")
```

we can view this data by using `ggplot2::geom_tile`

```{r, fig.height=8}
ptcld_validation_data %>% 
  dplyr::group_by(software, depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>%
  # collapse across study site
  dplyr::summarise(
    mean_f_score = mean(f_score, na.rm = T)
    # , med_f_score = median(f_score, na.rm = T)
    , sd_f_score = sd(f_score, na.rm = T)
    , n = dplyr::n()
  ) %>% 
  ggplot(mapping = aes(
    y = depth_maps_generation_quality
    , x = depth_maps_generation_filtering_mode
    , fill = mean_f_score
    , label = paste0(scales::comma(mean_f_score,accuracy = 0.01), "\n(n=", n,")")
  )) +
  geom_tile(color = "white") +
  geom_text(color = "white", size = 3) +
  facet_grid(cols = vars(software)) + 
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.9, direction = -1) +
  labs(
    x = "filtering mode"
    , y = "depth map quality"
    , fill = "F-score"
    , subtitle = "mean F-score and # of study sites"
  ) +
  theme_light() + 
  theme(
    legend.position = "none"
    , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
    , panel.background = element_blank()
    , panel.grid = element_blank()
    , plot.subtitle = element_text(hjust = 0.5)
    , strip.text = element_text(color = "black", face = "bold")
  )

ggplot2::ggsave("../data/fscore_comp_quick.png", height = 9, width = 8)
```


### Bayesian{#f_three_pred_mod_bays}

[Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/) describes the Hierarchical Bayesian approach to extend our two nominal predictor model to include another nominal predictor (referred to as "subject" here but the methodology applies for other nominal variables): 

> When every subject contributes many measurements to every cell, then the model of the situation is a straight-forward extension of the models we have already considered. We merely add "subject" as another nominal predictor in the model, with each individual subject being a level of the predictor. If there is one predictor other than subject, the model becomes
>
> $$ y = \beta_0 + \overrightarrow \beta_1 \overrightarrow x_1 + \overrightarrow \beta_S \overrightarrow x_S + \overrightarrow \beta_{1 \times S} \overrightarrow x_{1 \times S} $$
>
> This is exactly the two-predictor model we have already considered, with the second predictor being subject. When there are two predictors other than subject, the model becomes
> 
> \begin{align*}
> y = & \; \beta_0 & \text{baseline} \\
> & + \overrightarrow \beta_1 \overrightarrow x_1 + \overrightarrow \beta_2 \overrightarrow x_2 + \overrightarrow \beta_S \overrightarrow x_S  & \text{main effects} \\
> & + \overrightarrow \beta_{1 \times 2} \overrightarrow x_{1 \times 2} + \overrightarrow \beta_{1 \times S} \overrightarrow x_{1 \times S} + \overrightarrow \beta_{2 \times S} \overrightarrow x_{2 \times S} & \text{two-way interactions} \\
> & + \overrightarrow \beta_{1 \times 2 \times S} \overrightarrow x_{1 \times 2 \times S} & \text{three-way interactions}
> \end{align*}
> 
> This model includes all the two-way interactions of the factors, plus the three-way interaction. (p. 607)

The metric predicted variable with three nominal predictor variables model has the form:

\begin{align*}
y_{i} \sim & {\sf Normal} \bigl(\mu_{i}, \sigma_{y} \bigr) \\
\mu_{i} = & \beta_0 \\
& + \sum_{j} \beta_{1[j]} x_{1[j]} + \sum_{k} \beta_{2[k]} x_{2[k]} + \sum_{f} \beta_{3[f]} x_{3[f]}  \\
& + \sum_{j,k} \beta_{1\times2[j,k]} x_{1\times2[j,k]} + \sum_{j,f} \beta_{1\times3[j,f]} x_{1\times3[j,f]} + \sum_{k,f} \beta_{2\times3[k,f]} x_{2\times3[k,f]} \\
& + \sum_{j,k,f} \beta_{1\times2\times3[j,k,f]} x_{1\times2\times3[j,k,f]} \\
\beta_{0}  \sim & {\sf Normal} (0,100) \\ 
\beta_{1[j]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1}}) \\ 
\beta_{2[k]}  \sim & {\sf Normal} (0,\sigma_{\beta_{2}}) \\ 
\beta_{3[f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{3}}) \\ 
\beta_{1\times2[j,k]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1\times2}}) \\ 
\beta_{1\times3[j,f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1\times3}}) \\ 
\beta_{2\times3[k,f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{2\times3}}) \\ 
\sigma_{\beta_{1}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{2}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{1\times2}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{1\times3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{2\times3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{y} \sim & {\sf Cauchy} (0,109) \\ 
\end{align*}

, where $j$ is the depth map generation quality setting corresponding to observation $i$, $k$ is the depth map filtering mode setting corresponding to observation $i$, and $f$ is the processing software corresponding to observation $i$

*for this model, we'll define the priors following [Kurz](https://bookdown.org/content/3686/metric-predicted-variable-with-multiple-nominal-predictors.html#implementation-in-jags-brms.-1)* who notes that:

>The noise standard deviation $\sigma_y$ is depicted in the prior statement including the argument `class = sigma`...in order to be weakly informative, we will use the half-Cauchy. Recall that since the brms default is to set the lower bound for any variance parameter to 0, there’s no need to worry about doing so ourselves. So even though the syntax only indicates `cauchy`, it’s understood to mean Cauchy with a lower bound at zero; since the mean is usually 0, that makes this a half-Cauchy...The tails of the half-Cauchy are sufficiently fat that, in practice, I’ve found it doesn’t matter much what you set the $SD$ of its prior to.

```{r}
# from Kurz: 
gamma_a_b_from_omega_sigma <- function(mode, sd) {
  if (mode <= 0) stop("mode must be > 0")
  if (sd   <= 0) stop("sd must be > 0")
  rate <- (mode + sqrt(mode^2 + 4 * sd^2)) / (2 * sd^2)
  shape <- 1 + mode * rate
  return(list(shape = shape, rate = rate))
}

mean_y_temp <- mean(ptcld_validation_data$f_score)
sd_y_temp   <- sd(ptcld_validation_data$f_score)

omega_temp <- sd_y_temp / 2
sigma_temp <- 2 * sd_y_temp

s_r_temp <- gamma_a_b_from_omega_sigma(mode = omega_temp, sd = sigma_temp)

stanvars_temp <- 
  brms::stanvar(mean_y_temp,    name = "mean_y") + 
  brms::stanvar(sd_y_temp,      name = "sd_y") +
  brms::stanvar(s_r_temp$shape, name = "alpha") +
  brms::stanvar(s_r_temp$rate,  name = "beta")
```

Now fit the model.

```{r}
brms_f_mod3 = brms::brm(
  formula = f_score ~ 
    # baseline
    1 + 
    # main effects
    (1 | depth_maps_generation_quality) +
    (1 | depth_maps_generation_filtering_mode) + 
    (1 | software) +
    # two-way interactions
    (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode) +
    (1 | depth_maps_generation_quality:software) +
    (1 | depth_maps_generation_filtering_mode:software) +
    # three-way interactions
    (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode:software)
  , data = ptcld_validation_data
  , family = brms::brmsfamily(family = "gaussian")
  , iter = 50000, warmup = 25000, chains = 4
  , cores = round(parallel::detectCores()/2)
  , prior = c(
    brms::prior(normal(mean_y, sd_y * 5), class = "Intercept")
    , brms::prior(gamma(alpha, beta), class = "sd")
    , brms::prior(cauchy(0, sd_y), class = "sigma")
  )
  , stanvars = stanvars_temp
  , file = paste0(rootdir, "/fits/brms_f_mod3")
)
# https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
# https://mc-stan.org/misc/warnings.html#bulk-ess
# https://mc-stan.org/misc/warnings.html#tail-ess 
```

check the trace plots for problems with convergence of the Markov chains

```{r, fig.height=8}
plot(brms_f_mod3)
```

check the prior distributions

```{r}
# check priors
brms::prior_summary(brms_f_mod3) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling()
```

The `brms::brm` model summary

```{r}
brms_f_mod3 %>% 
  brms::posterior_summary() %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>%
  dplyr::rename_with(tolower) %>% 
  dplyr::filter(
    stringr::str_starts(parameter, "b_") 
    | stringr::str_starts(parameter, "r_") 
    | stringr::str_starts(parameter, "sd_") 
    | parameter == "sigma"
  ) %>%
  dplyr::mutate(
    parameter = parameter %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering")
  ) %>% 
  kableExtra::kbl(digits = 2, caption = "Bayesian 3 nominal predictors for F-score") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "8in")
```

We can look at the model noise standard deviation $\sigma_y$

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod3) %>% 
# plot
  ggplot(aes(x = sigma, y = 0)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21, point_size = 3
    , quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(latex2exp::TeX("$\\sigma_y$")) +
  theme_light()
```

how is it compared to our other models?

```{r}
# how is it compared to our other models?
dplyr::bind_rows(
    brms::as_draws_df(brms_f_mod1) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "one nominal predictor")
    , brms::as_draws_df(brms_f_mod2) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "two nominal predictor")
    , brms::as_draws_df(brms_f_mod3) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "three nominal predictor")
  ) %>% 
  dplyr::relocate(model) %>% 
  kableExtra::kbl(digits = 2, caption = "brms::brm model noise standard deviation comparison") %>% 
    kableExtra::kable_styling() %>% 
    kableExtra::scroll_box(width = "8in", height = "6in")
```

plot the posterior distributions of the conditional means with the median F-score and the 95% highest posterior density interval (HDI)

Note that how within `tidybayes::add_epred_draws`, we used the `re_formula` argument to average over the random effects of `software`. For this model we have to collapse across the software effects to compare the depth map quality and filtering mode setting effects.

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | depth_maps_generation_filtering_mode) + 
      (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  dplyr::mutate(depth_maps_generation_quality = depth_maps_generation_quality %>% forcats::fct_rev()) %>% 
  # plot
  ggplot(
    mapping = aes(
      y = value, x = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_eye(
      point_interval = median_hdi, .width = .95
      , slab_alpha = 0.8
      , interval_color = "black", linewidth = 1
      , point_color = "black", point_fill = "black", point_size = 1
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_y_continuous(breaks = scales::extended_breaks(n=10)) +
    facet_grid(cols = vars(depth_maps_generation_quality)) +
    labs(
      x = "filtering mode", y = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
      , fill = "Filtering Mode"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , legend.direction  = "horizontal"
      , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
      , strip.text = element_text(color = "black", face = "bold")
    ) 
    # guides(
    #   fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    # )
```

we can also make pairwise comparisons so long as we continue using `tidybayes::add_epred_draws` with the `re_formula` argument

```{r}
brms_contrast_temp =
  ptcld_validation_data %>%
    dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
    tidybayes::add_epred_draws(
      brms_f_mod3, allow_new_levels = T
      # this part is crucial
      , re_formula = ~ (1 | depth_maps_generation_quality) +
        (1 | depth_maps_generation_filtering_mode) + 
        (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode)
    ) %>% 
    dplyr::rename(value = .epred) %>% 
    tidybayes::compare_levels(
      value
      , by = depth_maps_generation_quality
      , comparison = 
        contrast_list
        # tidybayes::emmeans_comparison("revpairwise") 
        #"pairwise"
    ) %>% 
    dplyr::rename(contrast = depth_maps_generation_quality)

# separate contrast
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = contrast
    , delim = " - "
    , names = paste0(
        "sorter"
        , 1:(max(stringr::str_count(brms_contrast_temp$contrast, "-"))+1)
      )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = contrast %>% 
      forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
        as.numeric()
      )
    , depth_maps_generation_filtering_mode = depth_maps_generation_filtering_mode %>% 
      factor(
        levels = levels(ptcld_validation_data$depth_maps_generation_filtering_mode)
        , ordered = T
      )
  ) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast, depth_maps_generation_filtering_mode) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
    factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      , fill = sig_level # pct_gt_zero
      # , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = c(0.5,0.95)
      # , slab_fill = "gray22", slab_alpha = 1
      , interval_color = "black", point_color = "black", point_fill = "black"
      , justification = -0.01
    ) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
    scale_fill_viridis_d(
      option = "mako", begin = 0.3
      , drop = F
      # , labels = scales::percent
    ) +
    # scale_fill_viridis_c(
    #   option = "mako", begin = 0.3, direction = -1
    #   , labels = scales::percent
    # ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    facet_grid(cols = vars(depth_maps_generation_filtering_mode)) +
    labs(
      y = "depth map quality"
      , x = "constrast (f-score)"
      , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby filtering mode"
      , fill = "Pr(contrast > 0)"
    ) +
    theme_light() +
    theme(
      legend.position = "top"
      , legend.direction  = "horizontal"
      , strip.text = element_text(color = "black", face = "bold")
    ) +
    guides(
      fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    )
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, depth_maps_generation_filtering_mode) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, depth_maps_generation_filtering_mode) %>% 
  dplyr::select(contrast, depth_maps_generation_filtering_mode, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "quality contrast"
      , "filtering mode"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```

It might be more important to understand the difference in F-score by depth map quality and software rather than filtering mode since filtering mode had such a small effect on the SfM predictive ability

Note that how within `tidybayes::add_epred_draws`, we used the `re_formula` argument to average over the random effects of `depth_maps_generation_filtering_mode`. For this model we have to collapse across the filtering mode effects to compare the depth map quality and software setting effects.

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality, software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | software) + 
      (1 | depth_maps_generation_quality:software)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  dplyr::mutate(depth_maps_generation_quality = depth_maps_generation_quality %>% forcats::fct_rev()) %>% 
  # plot
  ggplot(
    mapping = aes(
      y = value, x = software
      , fill = software
    )
  ) +
  tidybayes::stat_eye(
    point_interval = median_hdi, .width = .95
    , slab_alpha = 0.8
    , interval_color = "black", linewidth = 1
    , point_color = "black", point_fill = "black", point_size = 1
  ) +
  scale_fill_viridis_d(option = "rocket", begin = 0.3, end = 0.9, drop = F) +
  scale_y_continuous(breaks = scales::extended_breaks(n=10)) +
  facet_grid(cols = vars(depth_maps_generation_quality)) +
  labs(
    x = "software", y = "f-score"
    , subtitle = "posterior distribution of conditional means with 95% HDI"
    , fill = "Filtering Mode"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
    , legend.direction  = "horizontal"
    , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
    , strip.text = element_text(color = "black", face = "bold")
  ) 
# guides(
#   fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
# )
```

we can also make pairwise comparisons so long as we continue using `tidybayes::add_epred_draws` with the `re_formula` argument

```{r}
brms_contrast_temp =
  ptcld_validation_data %>%
  dplyr::distinct(depth_maps_generation_quality, software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | software) + 
      (1 | depth_maps_generation_quality:software)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  tidybayes::compare_levels(
    value
    , by = depth_maps_generation_quality
    , comparison = 
      contrast_list[!stringr::str_detect(contrast_list, "lowest")]
      # contrast_list
  ) %>% 
  dplyr::rename(contrast = depth_maps_generation_quality)

# separate contrast
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = contrast
    , delim = " - "
    , names = paste0(
      "sorter"
      , 1:(max(stringr::str_count(brms_contrast_temp$contrast, "-"))+1)
    )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = contrast %>% 
      forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
          as.numeric()
      )
  ) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast, software) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
      factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      , fill = sig_level # pct_gt_zero
      # , fill = software
    )
  ) +
  tidybayes::stat_halfeye(
    point_interval = median_hdi, .width = c(0.5,0.95)
    # , slab_fill = "gray22", slab_alpha = 1
    , interval_color = "black", point_color = "black", point_fill = "black"
    , justification = -0.01
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
  scale_fill_viridis_d(
    option = "mako", begin = 0.3
    , drop = F
    # , labels = scales::percent
  ) +
  # scale_fill_viridis_c(
  #   option = "mako", begin = 0.3, direction = -1
  #   , labels = scales::percent
  # ) +
  scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
  facet_grid(cols = vars(software)) +
  labs(
    y = "depth map quality"
    , x = "constrast (f-score)"
    , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby software"
    , fill = "Pr(contrast > 0)"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , strip.text = element_text(color = "black", face = "bold")
  ) +
  guides(
    fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
  )
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, software) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, software) %>% 
  dplyr::select(contrast, software, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "quality contrast"
      , "software"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```

let's collapse across the filtering mode and software to compare the depth map quality setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_quality
      , fill = depth_maps_generation_quality
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "inferno", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "depth map quality", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [one nominal predictor model above](#f_one_pred_mod_bays) and [two nominal predictor model](#f_two_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod3") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod2
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_quality)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod2")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(brms_f_mod1) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod1")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(depth_maps_generation_quality), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[1:3]) +
  labs(
    y = "", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

let's also collapse across the depth map quality and software to compare the filtering mode setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "filtering mode", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [one nominal predictor model above](#f_one_pred_mod_bays) and [two nominal predictor model above](#f_two_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod3") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod2
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod2")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(depth_maps_generation_filtering_mode), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[2:3]) +
  labs(
    y = "filtering mode", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

to address one of our main questions, let's also collapse across the depth map quality and filtering mode setting to compare the software effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod3
    # this part is crucial
    , re_formula = ~ (1 | software)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = software
      , fill = software
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "rocket", begin = 0.3, end = 0.9, drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "filtering mode", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

Finally, we can quantify the variation in F-score by comparing the $\sigma$ posteriors.

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod3) %>% 
  dplyr::select(c(sigma,tidyselect::starts_with("sd_"))) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  # dplyr::group_by(name) %>% 
  # tidybayes::median_hdi(value) %>% 
  dplyr::mutate(
    name = name %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering") %>% 
      forcats::fct_reorder(value)
  ) %>%
# plot
  ggplot(aes(x = value, y = name)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21 #, point_size = 3
    , quantiles = 100
  ) +
  labs(x = "", y = "") +
  theme_light()
```

and perform model selection via information criteria with the `brms::loo_compare()` function

```{r}
brms_f_mod3 = brms::add_criterion(brms_f_mod3, criterion = c("loo", "waic"))
brms::loo_compare(brms_f_mod1, brms_f_mod2, brms_f_mod3, criterion = "loo")
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Three Nominal Predictors + site effects

Now, we'll add the average deflection from the baseline (i.e. the "grand mean") due to study site (i.e. the "subjects" in our data). The main effect for the study site will be added to our model with the combined influence of the depth map generation quality, the depth map filtering, and the software on the F-score.

In the model we use below, the study site is modeled as a "random effect." [Hobbs et al. (2024)](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1002/ecm.1598) describe a similar model:

>It is important to understand that fitting treatment intercepts and slopes as random rather than fixed means that our inference applied to all possible sites suitable for [inclusion in the study]. In contrast, assuming fixed effects of treatment would dramatically reduce the uncertainty about those effects, but would constrain inference to the four sites that we studied. (p. 13)

### Summary Statistics

Each study site contributes one observation per depth map quality, filtering mode, and software. That is, a row in the underlying data is unique by study site, software, depth map quality, and filtering mode.

```{r}
identical(
  # base data
  nrow(ptcld_validation_data)
  # distinct group
  , ptcld_validation_data %>% 
    dplyr::distinct(
      study_site
      , software
      , depth_maps_generation_quality
      , depth_maps_generation_filtering_mode
    ) %>% 
    nrow()
)
```

### Bayesian{#f_4_pred_mod_bays}

[Kruschke (2015)](https://sites.google.com/site/doingbayesiandataanalysis/) describes the Hierarchical Bayesian approach to describe groups of metric data with multiple nominal predictors when every subject ("study site" in our research) only contributes one observation per cell/condition: 

> \begin{align*}
> y = & \; \beta_0 \\
> & + \overrightarrow \beta_1 \overrightarrow x_1 + \overrightarrow \beta_2 \overrightarrow x_2 + \overrightarrow \beta_{1 \times 2} \overrightarrow x_{1 \times 2} \\
> & + \overrightarrow \beta_S \overrightarrow x_S
> \end{align*}

> In other words, we assume a main effect of subject, but no interaction of subject with other predictors. In this model, the subject effect (deflection) is constant across treatments, and the treatment effects (deflections) are constant across subjects. Notice that the model makes no requirement that every subject contributes a datum to every condition. Indeed, the model allows zero or multiple data per subject per condition. Bayesian estimation makes no assumptions or requirements that the design is balanced (i.e., has equal numbers of measurement in each cell). (p. 608)

and see section 20 from [Kurz's ebook supplement](https://bookdown.org/content/3686/metric-predicted-variable-with-multiple-nominal-predictors.html)

The metric predicted variable with three nominal predictor variables and subject-level effects model has the form:

\begin{align*}
y_{i} \sim & {\sf Normal} \bigl(\mu_{i}, \sigma_{y} \bigr) \\
\mu_{i} = & \beta_0 \\
& + \sum_{j} \beta_{1[j]} x_{1[j]} + \sum_{k} \beta_{2[k]} x_{2[k]} + \sum_{f} \beta_{3[f]} x_{3[f]} + \sum_{s} \beta_{4[s]} x_{4[s]}  \\
& + \sum_{j,k} \beta_{1\times2[j,k]} x_{1\times2[j,k]} + \sum_{j,f} \beta_{1\times3[j,f]} x_{1\times3[j,f]} + \sum_{k,f} \beta_{2\times3[k,f]} x_{2\times3[k,f]} \\
& + \sum_{j,k,f} \beta_{1\times2\times3[j,k,f]} x_{1\times2\times3[j,k,f]} \\
\beta_{0}  \sim & {\sf Normal} (0,100) \\ 
\beta_{1[j]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1}}) \\ 
\beta_{2[k]}  \sim & {\sf Normal} (0,\sigma_{\beta_{2}}) \\ 
\beta_{3[f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{3}}) \\ 
\beta_{4[s]}  \sim & {\sf Normal} (0,\sigma_{\beta_{4}}) \\ 
\beta_{1\times2[j,k]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1\times2}}) \\ 
\beta_{1\times3[j,f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1\times3}}) \\ 
\beta_{2\times3[k,f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{2\times3}}) \\ 
\sigma_{\beta_{1}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{2}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{4}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{1\times2}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{1\times3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{2\times3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{y} \sim & {\sf Cauchy} (0,109) \\ 
\end{align*}

, where $j$ is the depth map generation quality setting corresponding to observation $i$, $k$ is the depth map filtering mode setting corresponding to observation $i$, $f$ is the processing software corresponding to observation $i$, and $s$ is the study site corresponding to observation $i$

*for this model, we'll define the priors following [Kurz](https://bookdown.org/content/3686/metric-predicted-variable-with-multiple-nominal-predictors.html#implementation-in-jags-brms.-1)*:

```{r}
# from Kurz: 
gamma_a_b_from_omega_sigma <- function(mode, sd) {
  if (mode <= 0) stop("mode must be > 0")
  if (sd   <= 0) stop("sd must be > 0")
  rate <- (mode + sqrt(mode^2 + 4 * sd^2)) / (2 * sd^2)
  shape <- 1 + mode * rate
  return(list(shape = shape, rate = rate))
}

mean_y_temp <- mean(ptcld_validation_data$f_score)
sd_y_temp   <- sd(ptcld_validation_data$f_score)

omega_temp <- sd_y_temp / 2
sigma_temp <- 2 * sd_y_temp

s_r_temp <- gamma_a_b_from_omega_sigma(mode = omega_temp, sd = sigma_temp)

stanvars_temp <- 
  brms::stanvar(mean_y_temp,    name = "mean_y") + 
  brms::stanvar(sd_y_temp,      name = "sd_y") +
  brms::stanvar(s_r_temp$shape, name = "alpha") +
  brms::stanvar(s_r_temp$rate,  name = "beta")
```

Now fit the model.

```{r}
brms_f_mod4 = brms::brm(
  formula = f_score ~ 
    # baseline
    1 + 
    # main effects
    (1 | depth_maps_generation_quality) +
    (1 | depth_maps_generation_filtering_mode) + 
    (1 | software) +
    (1 | study_site) + # only fitting main effects of site and not interactions
    # two-way interactions
    (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode) +
    (1 | depth_maps_generation_quality:software) +
    (1 | depth_maps_generation_filtering_mode:software) +
    # three-way interactions
    (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode:software)
  , data = ptcld_validation_data
  , family = brms::brmsfamily(family = "gaussian")
  , iter = 50000, warmup = 25000, chains = 4
  , cores = round(parallel::detectCores()/2)
  , prior = c(
    brms::prior(normal(mean_y, sd_y * 5), class = "Intercept")
    , brms::prior(gamma(alpha, beta), class = "sd")
    , brms::prior(cauchy(0, sd_y), class = "sigma")
  )
  , stanvars = stanvars_temp
  # , control = list(adapt_delta = 0.999, max_treedepth = 13)
  , file = paste0(rootdir, "/fits/brms_f_mod4")
)
```

check the trace plots for problems with convergence of the Markov chains

```{r, fig.height=8}
plot(brms_f_mod4)
```

check the prior distributions

```{r}
# check priors
brms::prior_summary(brms_f_mod4) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling()
```

The `brms::brm` model summary

```{r}
brms_f_mod4 %>% 
  brms::posterior_summary() %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>%
  dplyr::rename_with(tolower) %>% 
  dplyr::filter(
    stringr::str_starts(parameter, "b_") 
    | stringr::str_starts(parameter, "r_") 
    | stringr::str_starts(parameter, "sd_") 
    | parameter == "sigma"
  ) %>%
  dplyr::mutate(
    parameter = parameter %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering")
  ) %>% 
  kableExtra::kbl(digits = 2, caption = "Bayesian 3 nominal predictors + study site effects for F-score") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```

We can look at the model noise standard deviation $\sigma_y$

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod4) %>% 
# plot
  ggplot(aes(x = sigma, y = 0)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21, point_size = 3
    , quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(latex2exp::TeX("$\\sigma_y$")) +
  theme_light()
```

how is it compared to our other models?

```{r}
# how is it compared to our other models?
dplyr::bind_rows(
    brms::as_draws_df(brms_f_mod1) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "one nominal predictor")
    , brms::as_draws_df(brms_f_mod2) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "two nominal predictor")
    , brms::as_draws_df(brms_f_mod3) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "three nominal predictor")
    , brms::as_draws_df(brms_f_mod4) %>% tidybayes::median_hdi(sigma) %>% dplyr::mutate(model = "three+site nominal predictor")
  ) %>% 
  dplyr::relocate(model) %>% 
  kableExtra::kbl(digits = 2, caption = "brms::brm model noise standard deviation comparison") %>% 
    kableExtra::kable_styling() %>% 
    kableExtra::scroll_box(width = "8in", height = "6in")
```

plot the posterior distributions of the conditional means with the median F-score and the 95% highest posterior density interval (HDI)

Note that how within `tidybayes::add_epred_draws`, we used the `re_formula` argument to average over the random effects of `study_site` (i.e., we left `(1 | study_site)` out of the formula) and `software` (i.e., we left `(1 | software)` and its interactions out of the formula).

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | depth_maps_generation_filtering_mode) + 
      (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  dplyr::mutate(depth_maps_generation_quality = depth_maps_generation_quality %>% forcats::fct_rev()) %>% 
  # plot
  ggplot(
    mapping = aes(
      y = value, x = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_eye(
      point_interval = median_hdi, .width = .95
      , slab_alpha = 0.8
      , interval_color = "black", linewidth = 1
      , point_color = "black", point_fill = "black", point_size = 1
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_y_continuous(breaks = scales::extended_breaks(n=10)) +
    facet_grid(cols = vars(depth_maps_generation_quality)) +
    labs(
      x = "filtering mode", y = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
      , fill = "Filtering Mode"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , legend.direction  = "horizontal"
      , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
      , strip.text = element_text(color = "black", face = "bold")
    ) 
    # guides(
    #   fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    # )
```

we can also make pairwise comparisons so long as we continue using `tidybayes::add_epred_draws` with the `re_formula` argument

```{r}
brms_contrast_temp =
  ptcld_validation_data %>%
    dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
    tidybayes::add_epred_draws(
      brms_f_mod4, allow_new_levels = T
      # this part is crucial
      , re_formula = ~ (1 | depth_maps_generation_quality) +
        (1 | depth_maps_generation_filtering_mode) + 
        (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode)
    ) %>% 
    dplyr::rename(value = .epred) %>% 
    tidybayes::compare_levels(
      value
      , by = depth_maps_generation_quality
      , comparison = 
        contrast_list
        # tidybayes::emmeans_comparison("revpairwise") 
        #"pairwise"
    ) %>% 
    dplyr::rename(contrast = depth_maps_generation_quality)

# separate contrast
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = contrast
    , delim = " - "
    , names = paste0(
        "sorter"
        , 1:(max(stringr::str_count(brms_contrast_temp$contrast, "-"))+1)
      )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = contrast %>% 
      forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
        as.numeric()
      )
    , depth_maps_generation_filtering_mode = depth_maps_generation_filtering_mode %>% 
      factor(
        levels = levels(ptcld_validation_data$depth_maps_generation_filtering_mode)
        , ordered = T
      )
  ) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast,depth_maps_generation_filtering_mode) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
    factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      , fill = sig_level # pct_gt_zero
      # , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = c(0.5,0.95)
      # , slab_fill = "gray22", slab_alpha = 1
      , interval_color = "black", point_color = "black", point_fill = "black"
      , justification = -0.01
    ) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
    scale_fill_viridis_d(
      option = "mako", begin = 0.3
      , drop = F
      # , labels = scales::percent
    ) +
    # scale_fill_viridis_c(
    #   option = "mako", begin = 0.3, direction = -1
    #   , labels = scales::percent
    # ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    facet_grid(cols = vars(depth_maps_generation_filtering_mode)) +
    labs(
      y = "depth map quality"
      , x = "constrast (f-score)"
      , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby filtering mode"
      , fill = "Pr(contrast > 0)"
    ) +
    theme_light() +
    theme(
      legend.position = "top"
      , legend.direction  = "horizontal"
      , strip.text = element_text(color = "black", face = "bold")
    ) +
    guides(
      fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    )
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, depth_maps_generation_filtering_mode) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, depth_maps_generation_filtering_mode) %>% 
  dplyr::select(contrast, depth_maps_generation_filtering_mode, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "quality contrast"
      , "filtering mode"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "8in")
```

It might be more important to understand the difference in F-score by depth map quality and software rather than filtering mode since filtering mode had such a small effect on the SfM predictive ability

Note that how within `tidybayes::add_epred_draws`, we used the `re_formula` argument to average over the random effects of `depth_maps_generation_filtering_mode`. For this model we have to collapse across the filtering mode effects to compare the depth map quality and software setting effects.

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality, software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | software) + 
      (1 | depth_maps_generation_quality:software)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  dplyr::mutate(depth_maps_generation_quality = depth_maps_generation_quality %>% forcats::fct_rev()) %>% 
  # plot
  ggplot(
    mapping = aes(
      y = value, x = software
      , fill = software
    )
  ) +
  tidybayes::stat_eye(
    point_interval = median_hdi, .width = .95
    , slab_alpha = 0.8
    , interval_color = "black", linewidth = 1
    , point_color = "black", point_fill = "black", point_size = 1
  ) +
  scale_fill_viridis_d(option = "rocket", begin = 0.3, end = 0.9, drop = F) +
  scale_y_continuous(breaks = scales::extended_breaks(n=10)) +
  facet_grid(cols = vars(depth_maps_generation_quality)) +
  labs(
    x = "software", y = "f-score"
    , subtitle = "posterior distribution of conditional means with 95% HDI"
    , fill = "Filtering Mode"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
    , legend.direction  = "horizontal"
    , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
    , strip.text = element_text(color = "black", face = "bold")
  ) 
# guides(
#   fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
# )
```

we can also make pairwise comparisons so long as we continue using `tidybayes::add_epred_draws` with the `re_formula` argument

```{r}
# get draws
qlty_sftwr_draws_temp = ptcld_validation_data %>%
  dplyr::distinct(depth_maps_generation_quality, software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | software) + 
      (1 | depth_maps_generation_quality:software)
  ) %>% 
  dplyr::rename(value = .epred)

# calculate contrast
brms_contrast_temp = qlty_sftwr_draws_temp %>% 
  tidybayes::compare_levels(
    value
    , by = depth_maps_generation_quality
    , comparison = 
      contrast_list[!stringr::str_detect(contrast_list, "lowest")]
      # contrast_list
  ) %>% 
  dplyr::rename(contrast = depth_maps_generation_quality)

# separate contrast
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = contrast
    , delim = " - "
    , names = paste0(
      "sorter"
      , 1:(max(stringr::str_count(brms_contrast_temp$contrast, "-"))+1)
    )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = contrast %>% 
      forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
          as.numeric()
      )
  ) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast, software) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
      factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      , fill = sig_level # pct_gt_zero
      # , fill = software
    )
  ) +
  tidybayes::stat_halfeye(
    point_interval = median_hdi, .width = c(0.5,0.95)
    # , slab_fill = "gray22", slab_alpha = 1
    , interval_color = "black", point_color = "black", point_fill = "black"
    , justification = -0.01
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
  scale_fill_viridis_d(
    option = "mako", begin = 0.3
    , drop = F
    # , labels = scales::percent
  ) +
  # scale_fill_viridis_c(
  #   option = "mako", begin = 0.3, direction = -1
  #   , labels = scales::percent
  # ) +
  scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
  facet_grid(cols = vars(software)) +
  labs(
    y = "depth map quality"
    , x = "constrast (f-score)"
    , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby software"
    , fill = "Pr(contrast > 0)"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , strip.text = element_text(color = "black", face = "bold")
  ) +
  guides(
    fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
  )
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, software) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, software) %>% 
  dplyr::select(contrast, software, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "quality contrast"
      , "software"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```

The contrasts above address the question "are there differences in F-score based on dense point cloud generation quality within each software?".

To address the different question of "are there differences in F-score based on the processing software used at a given dense point cloud generation quality?" we need to utilize a different formulation of the `comparison` parameter within our call to the `tidybayes::compare_levels` function and calculate the contrast by `software` instead

```{r}
# calculate contrast
brms_contrast_temp = qlty_sftwr_draws_temp %>% 
  tidybayes::compare_levels(
    value
    , by = software
    , comparison = "pairwise"
  ) %>% 
  dplyr::rename(contrast = software) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast, depth_maps_generation_quality) %>% 
  dplyr::mutate(
    is_gt_zero = abs(value) > 0.05
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
      factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      # , fill = sig_level
      , fill = pct_gt_zero
    )
  ) +
  tidybayes::stat_halfeye(
    point_interval = median_hdi, .width = c(0.5,0.95)
    # , slab_fill = "gray22", slab_alpha = 1
    , interval_color = "black", point_color = "black", point_fill = "black"
    , justification = -0.01
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
  scale_fill_viridis_c(
    option = "mako", begin = 0.3, direction = -1
    , breaks = scales::extended_breaks(n=6)
    , labels = scales::percent
  ) +
  scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
  facet_wrap(facets = vars(depth_maps_generation_quality), ncol = 2) +
  labs(
    y = "software"
    , x = "constrast (f-score)"
    , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby depth map quality"
    , fill = "Pr(abs(contrast) > 0.05)"
  ) +
  theme_light() +
  theme(
    legend.position = c(.75, .15)
    , strip.text = element_text(color = "black", face = "bold")
  ) 
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, depth_maps_generation_quality) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, depth_maps_generation_quality) %>% 
  dplyr::select(contrast, depth_maps_generation_quality, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "software contrast"
      , "depth map quality"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "8in")
```

let's collapse across the filtering mode, software, and study site to compare the depth map quality setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_quality
      , fill = depth_maps_generation_quality
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "inferno", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "depth map quality", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [one nominal predictor model above](#f_one_pred_mod_bays), [two nominal predictor model above](#f_two_pred_mod_bays), and [three nominal predictor model above](#f_three_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod4") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod3
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_quality)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod3")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod2
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_quality)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod2")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(brms_f_mod1) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod1")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(depth_maps_generation_quality), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[1:4]) +
  labs(
    y = "", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

let's also collapse across the depth map quality, software, and study site to compare the filtering mode setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "filtering mode", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [one nominal predictor model above](#f_one_pred_mod_bays), [two nominal predictor model above](#f_two_pred_mod_bays), and [three nominal predictor model above](#f_three_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod4") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod3
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod3")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod2
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod2")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(depth_maps_generation_filtering_mode), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[2:4]) +
  labs(
    y = "filtering mode", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

to address one of our main questions, let's also collapse across the depth map quality and filtering mode setting to compare the software effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4
    # this part is crucial
    , re_formula = ~ (1 | software)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = software
      , fill = software
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "rocket", begin = 0.3, end = 0.9, drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "filtering mode", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [three nominal predictor model above](#f_three_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod4
    # this part is crucial
    , re_formula = ~ (1 | software)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod4") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(software) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod3
        # this part is crucial
        , re_formula = ~ (1 | software)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod3")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(software), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[3:4]) +
  labs(
    y = "", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

Finally, we can quantify the variation in F-score by comparing the $\sigma$ posteriors.

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod4) %>% 
  dplyr::select(c(sigma,tidyselect::starts_with("sd_"))) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  # dplyr::group_by(name) %>% 
  # tidybayes::median_hdi(value) %>% 
  dplyr::mutate(
    name = name %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering") %>% 
      forcats::fct_reorder(value)
  ) %>%
# plot
  ggplot(aes(x = value, y = name)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21 #, point_size = 3
    , quantiles = 100
  ) +
  labs(x = "", y = "") +
  theme_light()
```

and perform model selection via information criteria with the `brms::loo_compare()` function

```{r}
brms_f_mod4 = brms::add_criterion(brms_f_mod4, criterion = c("loo", "waic"))
brms::loo_compare(brms_f_mod1, brms_f_mod2, brms_f_mod3, brms_f_mod4, criterion = "loo")
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```


## The beta: Three Nominal Predictors + site effects

To this point, we have been modelling F-score presuming a Gaussian likelihood. However, the beta likelihood more accurately represents the F-score data which is continuous and restricted within the range of $(0,1)$.

We borrow here from the excellent series on causal inference by [A. Solomon Kurz](https://solomonkurz.netlify.app/blog/2023-06-25-causal-inference-with-beta-regression/)

### Summary Statistics

let's check our underlying data for F-score (our dependent or $y$ variable)

```{r}
# distribution
ptcld_validation_data %>% 
  ggplot(mapping = aes(x = f_score)) + 
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = c(0,1)) +
  geom_density(fill = "lightblue", alpha = 0.7, color = NA) +
  labs(y="",x="f-score") +
  scale_y_continuous(breaks = c(0)) +
  scale_x_continuous(limits = c(0,1), breaks = scales::extended_breaks(10)) +
  theme_light() +
  theme(panel.grid = element_blank())
```

and the summary statistics

```{r}
ptcld_validation_data$f_score %>% 
  summary()
```

### Bayesian{#f_5_pred_mod_bays}

With the beta likelihood our model with three nominal predictor variables and subject-level effects the model becomes:

\begin{align*}
y_{i} \sim & {\sf Beta} \bigl(\mu_{i}, , \phi \bigr) \\
\operatorname{logit}(\mu_{i}) = & \beta_0 \\
& + \sum_{j} \beta_{1[j]} x_{1[j]} + \sum_{k} \beta_{2[k]} x_{2[k]} + \sum_{f} \beta_{3[f]} x_{3[f]} + \sum_{s} \beta_{4[s]} x_{4[s]}  \\
& + \sum_{j,k} \beta_{1\times2[j,k]} x_{1\times2[j,k]} + \sum_{j,f} \beta_{1\times3[j,f]} x_{1\times3[j,f]} + \sum_{k,f} \beta_{2\times3[k,f]} x_{2\times3[k,f]} \\
& + \sum_{j,k,f} \beta_{1\times2\times3[j,k,f]} x_{1\times2\times3[j,k,f]} \\
\beta_{0}  \sim & {\sf Normal} (0,1) \\ 
\beta_{1[j]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1}}) \\ 
\beta_{2[k]}  \sim & {\sf Normal} (0,\sigma_{\beta_{2}}) \\ 
\beta_{3[f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{3}}) \\ 
\beta_{4[s]}  \sim & {\sf Normal} (0,\sigma_{\beta_{4}}) \\ 
\beta_{1\times2[j,k]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1\times2}}) \\ 
\beta_{1\times3[j,f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{1\times3}}) \\ 
\beta_{2\times3[k,f]}  \sim & {\sf Normal} (0,\sigma_{\beta_{2\times3}}) \\ 
\sigma_{\beta_{1}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{2}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{4}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{1\times2}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{1\times3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\sigma_{\beta_{2\times3}} \sim & {\sf Gamma} (1.28,0.005) \\ 
\phi \sim & {\sf Gamma} (4,0.1) \\ 
\end{align*}

, where $j$ is the depth map generation quality setting corresponding to observation $i$, $k$ is the depth map filtering mode setting corresponding to observation $i$, $f$ is the processing software corresponding to observation $i$, and $s$ is the study site corresponding to observation $i$

Per [`brms`](https://cran.r-project.org/web/packages/brms/vignettes/brms_families.html), our $y$ is ${\sf Beta}$ distributed with the mean as $\mu$ and the concentration as $\phi$ which is sometimes called the *concentration*, *sample size* or *precision*

*Need to check our prior selection...just go with `brms` defaults for now*

Now fit the model.

```{r}
brms_f_mod5 = brms::brm(
  formula = f_score ~ 
    # baseline
    1 + 
    # main effects
    (1 | depth_maps_generation_quality) +
    (1 | depth_maps_generation_filtering_mode) + 
    (1 | software) +
    (1 | study_site) + # only fitting main effects of site and not interactions
    # two-way interactions
    (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode) +
    (1 | depth_maps_generation_quality:software) +
    (1 | depth_maps_generation_filtering_mode:software) +
    # three-way interactions
    (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode:software)
  , data = ptcld_validation_data
  , family = Beta(link = "logit")
  , iter = 50000, warmup = 25000, chains = 4
  , cores = round(parallel::detectCores()/2)
  , file = paste0(rootdir, "/fits/brms_f_mod5")
)
```

check the trace plots for problems with convergence of the Markov chains

```{r, fig.height=8}
plot(brms_f_mod5)
```

posterior-predictive check to make sure the model does an okay job simulating data that resemble the sample data

```{r}
# posterior predictive check
brms::pp_check(
    brms_f_mod5
    , type = "dens_overlay"
    , ndraws = 100
  ) + 
  labs(subtitle = "posterior-predictive check (overlaid densities)") +
  theme_light() +
  theme(legend.position = "top", legend.direction = "horizontal", legend.text = element_text(size = 14))
```

How’d we do capturing the conditional means and standard deviations by depth map generation quality?

```{r}
# means
p1_temp = brms::pp_check(
    brms_f_mod5
    , type = "stat_grouped" # "dens_overlay_grouped"
    , stat = "mean"
    , group = "depth_maps_generation_quality"
  ) + 
  scale_y_continuous(NULL, breaks = c(NULL)) +
  labs(subtitle = "means") +
  facet_grid(cols = vars(group), scales = "free") +
  theme_light()
# sds
p2_temp = brms::pp_check(
    brms_f_mod5
    , type = "stat_grouped" # "dens_overlay_grouped"
    , stat = "sd"
    , group = "depth_maps_generation_quality"
  ) + 
  scale_y_continuous(NULL, breaks = c(NULL)) +
  labs(subtitle = "sd's") +
  facet_grid(cols = vars(group), scales = "free") +
  theme_light()
# combine 
(p1_temp / p2_temp) &
  theme(legend.position = "none") &
  plot_annotation(
    title = "Posterior-predictive statistical checks\nby depth map quality"
    , subtitle = expression(
      "The dark blue lines are "*italic(T(y))*", and the light blue bars are for "*italic(T)(italic(y)[rep])*".")
  )

# brms_f_mod5 %>% 
#   brms::posterior_summary()
```

Both the means and sd's of the F-score are well represented across the different levels of depth map quality

check the prior distributions

```{r}
# check priors
brms::prior_summary(brms_f_mod5) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling()
```

The `brms::brm` model summary

```{r}
brms_f_mod5 %>% 
  brms::posterior_summary() %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>%
  dplyr::rename_with(tolower) %>% 
  dplyr::filter(
    stringr::str_starts(parameter, "b_") 
    | stringr::str_starts(parameter, "r_") 
    | stringr::str_starts(parameter, "sd_") 
    | parameter == "phi"
  ) %>%
  dplyr::mutate(
    parameter = parameter %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering")
  ) %>% 
  kableExtra::kbl(digits = 2, caption = "Bayesian two nominal predictors + study site effects for F-score") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "8in")
```

We can look at the model noise standard deviation (concentration) $\phi$

```{r}
# extract the posterior draws
brms::as_draws_df(brms_f_mod5) %>% 
# plot
  ggplot(aes(x = phi, y = 0)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21, point_size = 3
    , quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(latex2exp::TeX("$\\phi$")) +
  theme_light()

```

plot the posterior distributions of the conditional means with the median F-score and the 95% highest posterior density interval (HDI)

Note that how within `tidybayes::add_epred_draws`, we used the `re_formula` argument to average over the random effects of `study_site` (i.e., we left `(1 | study_site)` out of the formula) and `software` (i.e., we left `(1 | software)` and its interactions out of the formula).

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | depth_maps_generation_filtering_mode) + 
      (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  dplyr::mutate(depth_maps_generation_quality = depth_maps_generation_quality %>% forcats::fct_rev()) %>% 
  # plot
  ggplot(
    mapping = aes(
      y = value, x = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_eye(
      point_interval = median_hdi, .width = .95
      , slab_alpha = 0.8
      , interval_color = "black", linewidth = 1
      , point_color = "black", point_fill = "black", point_size = 1
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_y_continuous(breaks = scales::extended_breaks(n=10)) +
    facet_grid(cols = vars(depth_maps_generation_quality)) +
    labs(
      x = "filtering mode", y = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
      , fill = "Filtering Mode"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , legend.direction  = "horizontal"
      , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
      , strip.text = element_text(color = "black", face = "bold")
    ) 
    # guides(
    #   fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    # )
```

we can also make pairwise comparisons so long as we continue using `tidybayes::add_epred_draws` with the `re_formula` argument

```{r}
brms_contrast_temp =
  ptcld_validation_data %>%
    dplyr::distinct(depth_maps_generation_quality, depth_maps_generation_filtering_mode) %>% 
    tidybayes::add_epred_draws(
      brms_f_mod5, allow_new_levels = T
      # this part is crucial
      , re_formula = ~ (1 | depth_maps_generation_quality) +
        (1 | depth_maps_generation_filtering_mode) + 
        (1 | depth_maps_generation_quality:depth_maps_generation_filtering_mode)
    ) %>% 
    dplyr::rename(value = .epred) %>% 
    tidybayes::compare_levels(
      value
      , by = depth_maps_generation_quality
      , comparison = 
        contrast_list
        # tidybayes::emmeans_comparison("revpairwise") 
        #"pairwise"
    ) %>% 
    dplyr::rename(contrast = depth_maps_generation_quality)

# separate contrast
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = contrast
    , delim = " - "
    , names = paste0(
        "sorter"
        , 1:(max(stringr::str_count(brms_contrast_temp$contrast, "-"))+1)
      )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = contrast %>% 
      forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
        as.numeric()
      )
    , depth_maps_generation_filtering_mode = depth_maps_generation_filtering_mode %>% 
      factor(
        levels = levels(ptcld_validation_data$depth_maps_generation_filtering_mode)
        , ordered = T
      )
  ) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast,depth_maps_generation_filtering_mode) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
    factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      , fill = sig_level # pct_gt_zero
      # , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = c(0.5,0.95)
      # , slab_fill = "gray22", slab_alpha = 1
      , interval_color = "black", point_color = "black", point_fill = "black"
      , justification = -0.01
    ) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
    scale_fill_viridis_d(
      option = "mako", begin = 0.3
      , drop = F
      # , labels = scales::percent
    ) +
    # scale_fill_viridis_c(
    #   option = "mako", begin = 0.3, direction = -1
    #   , labels = scales::percent
    # ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    facet_grid(cols = vars(depth_maps_generation_filtering_mode)) +
    labs(
      y = "depth map quality"
      , x = "constrast (f-score)"
      , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby filtering mode"
      , fill = "Pr(contrast > 0)"
    ) +
    theme_light() +
    theme(
      legend.position = "top"
      , legend.direction  = "horizontal"
      , strip.text = element_text(color = "black", face = "bold")
    ) +
    guides(
      fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
    )
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, depth_maps_generation_filtering_mode) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, depth_maps_generation_filtering_mode) %>% 
  dplyr::select(contrast, depth_maps_generation_filtering_mode, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "quality contrast"
      , "filtering mode"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "8in")
```

It might be more important to understand the difference in F-score by depth map quality and software rather than filtering mode since filtering mode had such a small effect on the SfM predictive ability

Note that how within `tidybayes::add_epred_draws`, we used the `re_formula` argument to average over the random effects of `depth_maps_generation_filtering_mode`. For this model we have to collapse across the filtering mode effects to compare the depth map quality and software setting effects.

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality, software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | software) + 
      (1 | depth_maps_generation_quality:software)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  dplyr::mutate(depth_maps_generation_quality = depth_maps_generation_quality %>% forcats::fct_rev()) %>% 
  # plot
  ggplot(
    mapping = aes(
      y = value, x = software
      , fill = software
    )
  ) +
  tidybayes::stat_eye(
    point_interval = median_hdi, .width = .95
    , slab_alpha = 0.8
    , interval_color = "black", linewidth = 1
    , point_color = "black", point_fill = "black", point_size = 1
  ) +
  scale_fill_viridis_d(option = "rocket", begin = 0.3, end = 0.9, drop = F) +
  scale_y_continuous(breaks = scales::extended_breaks(n=10)) +
  facet_grid(cols = vars(depth_maps_generation_quality)) +
  labs(
    x = "software", y = "f-score"
    , subtitle = "posterior distribution of conditional means with 95% HDI"
    , fill = "Filtering Mode"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
    , legend.direction  = "horizontal"
    , axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
    , strip.text = element_text(color = "black", face = "bold")
  ) 
# guides(
#   fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
# )
```

we can also make pairwise comparisons so long as we continue using `tidybayes::add_epred_draws` with the `re_formula` argument

```{r}
# get draws
qlty_sftwr_draws_temp = ptcld_validation_data %>%
  dplyr::distinct(depth_maps_generation_quality, software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5, allow_new_levels = T
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality) +
      (1 | software) + 
      (1 | depth_maps_generation_quality:software)
  ) %>% 
  dplyr::rename(value = .epred)

# calculate contrast
brms_contrast_temp = qlty_sftwr_draws_temp %>% 
  tidybayes::compare_levels(
    value
    , by = depth_maps_generation_quality
    , comparison = 
      contrast_list[!stringr::str_detect(contrast_list, "lowest")]
      # contrast_list
  ) %>% 
  dplyr::rename(contrast = depth_maps_generation_quality)

# separate contrast
brms_contrast_temp = brms_contrast_temp %>% 
  dplyr::ungroup() %>% 
  tidyr::separate_wider_delim(
    cols = contrast
    , delim = " - "
    , names = paste0(
      "sorter"
      , 1:(max(stringr::str_count(brms_contrast_temp$contrast, "-"))+1)
    )
    , too_few = "align_start"
    , cols_remove = F
  ) %>% 
  dplyr::filter(sorter1!=sorter2) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("sorter")
      , .fns = function(x){factor(
        x, ordered = T
        , levels = levels(ptcld_validation_data$depth_maps_generation_quality)
      )}
    )
    , contrast = contrast %>% 
      forcats::fct_reorder(
        paste0(as.numeric(sorter1), as.numeric(sorter2)) %>% 
          as.numeric()
      )
  ) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast, software) %>% 
  dplyr::mutate(
    is_gt_zero = value > 0
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
      factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      , fill = sig_level # pct_gt_zero
      # , fill = software
    )
  ) +
  tidybayes::stat_halfeye(
    point_interval = median_hdi, .width = c(0.5,0.95)
    # , slab_fill = "gray22", slab_alpha = 1
    , interval_color = "black", point_color = "black", point_fill = "black"
    , justification = -0.01
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
  scale_fill_viridis_d(
    option = "mako", begin = 0.3
    , drop = F
    # , labels = scales::percent
  ) +
  # scale_fill_viridis_c(
  #   option = "mako", begin = 0.3, direction = -1
  #   , labels = scales::percent
  # ) +
  scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
  facet_grid(cols = vars(software)) +
  labs(
    y = "depth map quality"
    , x = "constrast (f-score)"
    , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby software"
    , fill = "Pr(contrast > 0)"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , strip.text = element_text(color = "black", face = "bold")
  ) +
  guides(
    fill = guide_legend(reverse = T, override.aes = list(alpha = 1, color = NA, shape = NA, lwd = NA))
  )
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, software) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, software) %>% 
  dplyr::select(contrast, software, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "quality contrast"
      , "software"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "6in")
```

The contrasts above address the question "are there differences in F-score based on dense point cloud generation quality within each software?".

To address the different question of "are there differences in F-score based on the processing software used at a given dense point cloud generation quality?" we need to utilize a different formulation of the `comparison` parameter within our call to the `tidybayes::compare_levels` function and calculate the contrast by `software` instead

```{r}
# calculate contrast
brms_contrast_temp = qlty_sftwr_draws_temp %>% 
  tidybayes::compare_levels(
    value
    , by = software
    , comparison = "pairwise"
  ) %>% 
  dplyr::rename(contrast = software) %>% 
  # median_hdi summary for coloring 
  dplyr::group_by(contrast, depth_maps_generation_quality) %>% 
  dplyr::mutate(
    is_gt_zero = abs(value) > 0.05
    , pct_gt_zero = sum(is_gt_zero)/dplyr::n()
    , sig_level = dplyr::case_when(
      pct_gt_zero > 0.99 ~ 0
      , pct_gt_zero > 0.95 ~ 1
      , pct_gt_zero > 0.9 ~ 2
      , pct_gt_zero > 0.8 ~ 3
      , T ~ 4
    ) %>% 
      factor(levels = c(0:4), labels = c(">99%","95%","90%","80%","<80%"), ordered = T)
  )

# what?
brms_contrast_temp %>% dplyr::glimpse()
```

plot it

```{r}
# plot it
brms_contrast_temp %>% 
  ggplot(
    mapping = aes(
      x = value, y = contrast
      # , fill = sig_level
      , fill = pct_gt_zero
    )
  ) +
  tidybayes::stat_halfeye(
    point_interval = median_hdi, .width = c(0.5,0.95)
    # , slab_fill = "gray22", slab_alpha = 1
    , interval_color = "black", point_color = "black", point_fill = "black"
    , justification = -0.01
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray44") +
  scale_fill_viridis_c(
    option = "mako", begin = 0.3, direction = -1
    , breaks = scales::extended_breaks(n=6)
    , labels = scales::percent
  ) +
  scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
  facet_wrap(facets = vars(depth_maps_generation_quality), ncol = 2) +
  labs(
    y = "software"
    , x = "constrast (f-score)"
    , subtitle = "95% & 50% HDI of the posterior distribution of conditional mean group constrasts\nby depth map quality"
    , fill = "Pr(abs(contrast) > 0.05)"
  ) +
  theme_light() +
  theme(
    legend.position = c(.75, .15)
    , strip.text = element_text(color = "black", face = "bold")
  ) 
```

and summarize these contrasts

```{r}
brms_contrast_temp %>%
  dplyr::group_by(contrast, depth_maps_generation_quality) %>% 
  tidybayes::median_hdi(value) %>% 
  dplyr::arrange(contrast, depth_maps_generation_quality) %>% 
  dplyr::select(contrast, depth_maps_generation_quality, value, .lower, .upper) %>% 
  dplyr::rename(difference=value) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "brms::brm model: 95% HDI of the posterior distribution of conditional mean group constrasts"
    , col.names = c(
      "software contrast"
      , "depth map quality"
      , "difference (f-score)"
      , "conf.low", "conf.high"
    )
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "8in", height = "8in")
```

let's collapse across the filtering mode, software, and study site to compare the depth map quality setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_quality
      , fill = depth_maps_generation_quality
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "inferno", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "depth map quality", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [one nominal predictor model above](#f_one_pred_mod_bays), [two nominal predictor model above](#f_two_pred_mod_bays), [three nominal predictor model above](#f_three_pred_mod_bays), and [three nominal predictor + site effects model above](#f_4_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_quality) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_quality)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod5") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod4
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_quality)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod4")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod3
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_quality)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod3")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod2
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_quality)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod2")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_quality) %>% 
      tidybayes::add_epred_draws(brms_f_mod1) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod1")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(depth_maps_generation_quality), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  # scale_color_viridis_d(option = "turbo", begin = 0.2, end = 0.8) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)) +
  labs(
    y = "", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

let's also collapse across the depth map quality, software, and study site to compare the filtering mode setting effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = depth_maps_generation_filtering_mode
      , fill = depth_maps_generation_filtering_mode
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "plasma", drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "filtering mode", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [one nominal predictor model above](#f_one_pred_mod_bays), [two nominal predictor model above](#f_two_pred_mod_bays), and [three nominal predictor model above](#f_three_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5
    # this part is crucial
    , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod5") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod4
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod4")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod3
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod3")
    , ptcld_validation_data %>% 
      dplyr::distinct(depth_maps_generation_filtering_mode) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod2
        # this part is crucial
        , re_formula = ~ (1 | depth_maps_generation_filtering_mode)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod2")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(depth_maps_generation_filtering_mode), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[2:5]) +
  labs(
    y = "filtering mode", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

to address one of our main questions, let's also collapse across the depth map quality and filtering mode setting to compare the software effect.
In a hierarchical model structure, we have to make use of the `re_formula` argument within `tidybayes::add_epred_draws`

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5
    # this part is crucial
    , re_formula = ~ (1 | software)
  ) %>% 
  dplyr::rename(value = .epred) %>% 
  # plot
  ggplot(
    mapping = aes(
      x = value, y = software
      , fill = software
    )
  ) +
    tidybayes::stat_halfeye(
      point_interval = median_hdi, .width = .95
      , interval_color = "gray66"
      , shape = 21, point_color = "gray66", point_fill = "black"
      , justification = -0.01
    ) +
    scale_fill_viridis_d(option = "rocket", begin = 0.3, end = 0.9, drop = F) +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      y = "filtering mode", x = "f-score"
      , subtitle = "posterior distribution of conditional means with 95% HDI"
    ) +
    theme_light() +
    theme(legend.position = "none")
```

let's compare these results to the results from our [three nominal predictor model above](#f_three_pred_mod_bays) and and [three nominal predictor + site effects model above](#f_4_pred_mod_bays)

```{r}
ptcld_validation_data %>% 
  dplyr::distinct(software) %>% 
  tidybayes::add_epred_draws(
    brms_f_mod5
    # this part is crucial
    , re_formula = ~ (1 | software)
  ) %>% 
  dplyr::mutate(value = .epred, src = "brms_f_mod5") %>%
  dplyr::bind_rows(
    ptcld_validation_data %>% 
      dplyr::distinct(software) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod4
        # this part is crucial
        , re_formula = ~ (1 | software)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod4")
    , ptcld_validation_data %>% 
      dplyr::distinct(software) %>% 
      tidybayes::add_epred_draws(
        brms_f_mod3
        # this part is crucial
        , re_formula = ~ (1 | software)
      ) %>% 
      dplyr::mutate(value = .epred, src = "brms_f_mod3")
  ) %>% 
  ggplot(mapping = aes(y = src, x = value, color = src, group = src)) +
  tidybayes::stat_pointinterval(position = "dodge") +
  facet_grid(rows = vars(software), switch = "y") +
  scale_y_discrete(NULL, breaks = NULL) +
  scale_color_manual(values = viridis::turbo(n = 5, begin = 0.2, end = 0.8)[3:5]) +
  labs(
    y = "", x = "f-score"
    , color = "model"
  ) +
  theme_light() +
  theme(legend.position = "top", strip.text = element_text(color = "black", face = "bold"))
  
```

```{r, include=FALSE, eval=FALSE}
# Finally, we can quantify the variation in F-score by comparing the $\sigma$ posteriors.
# extract the posterior draws
brms::as_draws_df(brms_f_mod5) %>% 
  dplyr::select(c(phi,tidyselect::starts_with("sd_"))) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  # dplyr::group_by(name) %>% 
  # tidybayes::median_hdi(value) %>% 
  dplyr::mutate(
    name = name %>% 
      stringr::str_replace_all("depth_maps_generation_quality", "quality") %>% 
      stringr::str_replace_all("depth_maps_generation_filtering_mode", "filtering") %>% 
      forcats::fct_reorder(value)
    , value = plogis(value)
  ) %>%
# plot
  ggplot(aes(x = value, y = name)) +
  tidybayes::stat_dotsinterval(
    point_interval = median_hdi, .width = .95
    , justification = -0.04
    , shape = 21 #, point_size = 3
    , quantiles = 100
  ) +
  labs(x = "", y = "") +
  scale_x_continuous(limits = c(-0.1,1.1)) +
  theme_light()
```

and perform model selection via information criteria with the `brms::loo_compare()` function

```{r}
brms_f_mod5 = brms::add_criterion(brms_f_mod5, criterion = c("loo", "waic"))
brms::loo_compare(brms_f_mod1, brms_f_mod2, brms_f_mod3, brms_f_mod4, brms_f_mod5, criterion = "loo")
# brms::model_weights(brms_f_mod1, brms_f_mod2, brms_f_mod3, brms_f_mod4) %>% round(3)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

```{r one-sum-norm, eval=FALSE, include=FALSE}
brms_f_mod1 %>% 
  plot(variable = c("^b_", "sigma"), regex = TRUE)

brms::as_draws_df(
        brms_f_mod1
        , variable = c("^b_", "shape")
        , regex = TRUE
      ) %>% 
      # quick way to get a table of summary statistics and diagnostics
      posterior::summarize_draws(
        "mean", "median", "sd"
        ,  ~quantile(.x, probs = c(
          0.05, 0.95
          # , 0.025, 0.975
        ))
        , "rhat"
      )


# Check normality assumption by groups. Computing Shapiro-Wilk test for each group level. If the data is normally distributed, the p-value should be greater than 0.05.


ptcld_validation_data %>% 
  dplyr::group_by(depth_maps_generation_quality) %>% 
  rstatix::shapiro_test(f_score) %>% 
  kableExtra::kbl(digits = 3) %>% 
  kableExtra::kable_styling()

# The time is not normally distributed (p < 0.05) for most groups as assessed by Shapiro-Wilk’s test of normality. 
# 
# If you have doubt about the normality of the data, you can use the Kruskal-Wallis test, which is the non-parametric alternative to one-way ANOVA test.
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```
