# Overstory Field Validation Data{#ostory_valid}

[Tinkham and Swayze (2021; p.6)](https://scholar.google.com/scholar?oi=bibs&hl=en&cluster=11260597505702247290) describe a methodology for matching UAS *detected* trees with stem mapped trees identified via traditional field *survey* methods. Note, *detected* trees in the excerpt below references UAS detected trees while *survey* trees references field-based stem mapped trees:

*Each of the detected tree outputs was matched with survey tree locations through an iterative process. Iteratively, a detected tree was selected, and all survey trees within a 3 m radius and 10% height of the detected tree were identified. If a survey tree met both the location and height precision requirements, it was considered a true positive (TP) detection, and both the survey and detected trees were removed from further matching. However, if no match was made, the detected tree was considered a commission (Co) and removed from further matching. This process was repeated until all detected trees were classified as true positive or commission, with all unmatched survey trees classified as omission (Om). Overall tree detection performance was described using the F-score metric.*

The F-score incorporates true positive, commission, and omission rates to determine how well the UAS detected trees represent the field-based stem mapped trees. As a measure of predictive performance, the highest possible value of an F-score is 1.0, indicating perfect precision and recall, and the lowest possible value is 0, if either precision or recall are zero.

$$
\textrm{F-score} = 2 \times \frac{\bigl(\frac{TP}{TP+Om} \times \frac{TP}{TP+Co} \bigr)}{\bigl(\frac{TP}{TP+Om} + \frac{TP}{TP+Co} \bigr)}
$$

The process to match UAS detected trees to field stem mapped trees implemented here is slightly different than the process described above. To match the data parametrization from the UAS point cloud processing workflow, only stem-mapped trees above `r min(chm_rast %>% terra::values(), na.rm = T) %>% round(1)` m were considered for analysis. Each UAS detected tree was matched with stem-mapped tree locations that were within a 3 m radius and 2 m height of the UAS detected tree. The matched UAS and stem-mapped tree pairs were jointly compared (rather than iteratively) to select the pair that minimized the height difference for both the stem-mapped tree and the UAS detected tree to ensure that only one UAS detected tree was selected for each stem-mapped tree. If more than one UAS detected tree had the same height difference to a stem-mapped tree, the UAS detected tree spatially nearest to the stem-mapped tree was selected as the match. These UAS detected trees with a paired stem-mapped tree after this filtering process were considered true positive ($TP$) detections.

To determine UAS detected tree commissions (i.e. UAS detected trees within the overstory plot for which there was no stem-mapped tree pair; $Co$) this analysis used the 2023-06 BHEF overstory field survey plot center and plot radius with a minimum DBH of 5 in (12.69 cm) as only trees above this size were sampled as part of the overstory survey. UAS detected trees within this radius with an [estimated DBH](#local_mod) over 5 in (12.69 cm) that did not have a matched stem-mapped tree pair were considered commissions ($Co$). The 2023-06 BHEF field surveys used $\frac{1}{10}$ acre (404.686 m^2^) plots with a 37.24 ft (11.35 m) radius for overstory sampling and $\frac{1}{400}$ acre (10.117 m^2^) plots with a 5.89 ft (1.795 m) radius for regeneration sampling. All unmatched stem-mapped survey trees were classified as omissions ($Om$).

## Load Field Data

Load the overstory data, set parameters for processing, and map

```{r fld-dta-ld, results='hide'}
##################################################################################################
## USER DEFINED PARAMETERS
##################################################################################################
### Only include trees above a specific height 
  # e.g., set to match the minimum height used in point cloud processing
min_tree_height_m = min(chm_rast %>% terra::values(), na.rm = T)
### Only include trees above a specific DBH
  # e.g., set to match the minimum tree DBH for field sampling of overstory trees
min_tree_dbh_cm = 5*2.54 # 3 in = 7.62 cm
### Find all nearest neighbors within maximum distance and then filter them based on 
### distance and height error 
max_dist_m = 3
max_height_error_m = 2
##################################################################################################
## USER DEFINED PARAMETERS
##################################################################################################
# load Field Validation Data
field_overstory = readr::read_csv(
    file = "../data/field_data/Voodoo_overstory_data.csv"
  ) %>% 
  dplyr::rename_with(~ .x %>%
      # replace all non alphanumeric with _
      stringr::str_replace_all("[^[:alnum:] ]+", "_") %>%
      # remove any _ at the end of the string
      stringr::str_remove("[_]$") %>% 
      tolower()
  ) %>% 
  dplyr::filter(
    !is.na(tree_diameter_inches)
    & !is.na(tree_height_feet)
    & !is.na(tree_utm_x)
    & !is.na(tree_utm_y)
  ) %>% 
  dplyr::mutate(
    field_dbh_cm = tree_diameter_inches/0.394
    , field_tree_height_m = tree_height_feet/3.281
    , field_plottree_id = paste0(
      stringr::str_pad(plot_id, width = 4, side = "left", pad = "0")
      , "-"
      , stringr::str_pad(tree_id, width = 4, side = "left", pad = "0")
    )
  ) %>% 
  dplyr::filter(
    field_tree_height_m >= min_tree_height_m
    & field_dbh_cm >= min_tree_dbh_cm
  ) %>% 
  dplyr::rename(
    field_dbh_in = tree_diameter_inches
    , field_tree_height_ft = tree_height_feet
  ) %>% 
  sf::st_as_sf(
    coords = c("tree_utm_x", "tree_utm_y")
    , crs = sf::st_crs(treetops_sf_with_dbh)
    , remove=F
  ) %>% 
  # keep only tree points that are within the uas flight boundary
  sf::st_intersection(las_ctg_dta)

  # field_overstory %>% 
  #   sf::st_drop_geometry() %>% 
  #   dplyr::summarize(n = dplyr::n(), unique_tree = dplyr::n_distinct(field_plottree_id))
  
```

```{r fld-dta-str}
field_overstory %>% 
  dplyr::glimpse()
```

Overstory field data map

```{r fld-dta-map, results='asis', fig.height = 5}
field_overstory %>% 
  sf::st_drop_geometry() %>% 
  dplyr::distinct(plot_id, utm_x, utm_y) %>% 
  sf::st_as_sf(coords = c("utm_x", "utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)) %>% 
  sf::st_buffer(11.35) %>% 
  dplyr::mutate(x=1) %>% 
mapview::mapview(col.regions = "blue", layer.name = "plot", alpha.regions = 0.6) +
mapview::mapview(field_overstory, col.regions = "orange", cex = 2, layer.name = "trees") +
mapview::mapview(
  harvests
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
)
```

## Build UAS-Field Validation Data

### True Positive Identification

The UAS detected and stem-mapped tree pairs identified in this filtering process (detailed above) were considered true positive ($TP$) detections.

```{r fld-dta-tp, results='hide'}
## BUFFER THE UAS TREES AND SPATIALLY MATCH FIELD TREES BASED ON THAT BUFFER   
## get FIELD trees within radius OF UAS TREES
  potential_tree_pairs_temp = treetops_sf_with_dbh %>% 
    dplyr::select(treeID, tree_height_m) %>% 
    # buffer point
    sf::st_buffer(max_dist_m) %>% 
    # spatial join with all FIELD tree points
    sf::st_join(
      field_overstory %>% 
        dplyr::filter(
          field_tree_height_m >= min_tree_height_m
          & field_dbh_cm >= min_tree_dbh_cm
        ) %>% 
        dplyr::select(
          field_plottree_id, field_tree_height_m
          , tree_utm_x, tree_utm_y
        )
      , join = st_intersects
      , left = F # performs inner join to only keep uas trees with a match
    ) %>% 
    # calculate height difference
    dplyr::mutate(
      height_diff_m = abs(tree_height_m-field_tree_height_m)
      , height_diff_pct = height_diff_m/field_tree_height_m
    ) %>% 
    # removes tree pairs that are outside of the allowable error
    # dplyr::filter(height_diff_pct <= max_height_error_pct) %>% 
    dplyr::filter(height_diff_m <= max_height_error_m) %>% 
    dplyr::select(-c(height_diff_m)) %>% 
    dplyr::relocate(treeID, field_plottree_id)

## apply pair selection criteria if there are potential tree pairs
  if(nrow(potential_tree_pairs_temp)>0){
    ## calculate row by row distances and height differences
      potential_tree_pairs_temp = potential_tree_pairs_temp %>% 
        # this is the position of the uas tree
        sf::st_centroid() %>% 
        sf::st_set_geometry("geom1") %>% 
        dplyr::bind_cols(
          potential_tree_pairs_temp %>% 
            sf::st_drop_geometry() %>% 
            dplyr::select("tree_utm_x", "tree_utm_y") %>% 
            # this is the position of the field tree
            sf::st_as_sf(
              coords = c("tree_utm_x", "tree_utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)
            ) %>% 
            sf::st_set_geometry("geom2")
        ) %>% 
        dplyr::mutate(
          distance_m = sf::st_distance(geom1, geom2, by_element = T) %>% as.numeric()
        ) %>% 
        sf::st_drop_geometry() %>% 
        dplyr::select(-c(tree_utm_x, tree_utm_y, geom2))
    
    ## define function to select the best tree pair
      select_best_tree_pair_fn <- function(df) {
       df %>% 
        dplyr::group_by(field_plottree_id) %>% 
        dplyr::arrange(field_plottree_id, height_diff_pct, distance_m, desc(tree_height_m), treeID) %>% 
        dplyr::mutate(
          # at the field tree level...the number of uas trees
          n_uas_trees = dplyr::n()
          # at the field tree level...
            # the closest uas tree in height tie breaker distance, uas_tree_height_m, id
          , rank_within_field_tree = dplyr::row_number()
        ) %>% 
        dplyr::group_by(treeID) %>% 
        dplyr::arrange(treeID, height_diff_pct, distance_m, desc(field_tree_height_m), field_plottree_id) %>% 
        dplyr::mutate(
          # at the uas tree level...the number of field trees
          n_field_trees = dplyr::n()
          # at the field tree level...
            # the closest field tree in height tie breaker distance, uas_tree_height_m, id
          , rank_within_uas_tree = dplyr::row_number()
        ) %>% 
        dplyr::ungroup() %>% 
        # select the uas-field tree pair with the minimum height difference
        dplyr::filter(
          rank_within_field_tree == 1
          & rank_within_uas_tree == 1
        ) %>% 
        # remove columns
        dplyr::select(
          -c(tidyselect::starts_with("rank_"), tidyselect::starts_with("n_"))
        )
      }
    
    ## first filter for tree pairs
      true_positive_trees = select_best_tree_pair_fn(potential_tree_pairs_temp)
    
    ##remove matches from potential tree pairs 
      potential_tree_pairs_temp = potential_tree_pairs_temp %>% 
        dplyr::filter(
          !(treeID %in% true_positive_trees$treeID)
          & !(field_plottree_id %in% true_positive_trees$field_plottree_id)
        )
      
    ## keep filtering for best pair until no unique pairs remain
      while(nrow(potential_tree_pairs_temp)>0) {
        # keep filtering for best pair until no unique pairs remain
        true_positive_trees = true_positive_trees %>% 
          dplyr::bind_rows(
            select_best_tree_pair_fn(potential_tree_pairs_temp)
          )
        #remove matches from potential tree pairs 
        potential_tree_pairs_temp = potential_tree_pairs_temp %>% 
          dplyr::filter(
            !(treeID %in% true_positive_trees$treeID)
            & !(field_plottree_id %in% true_positive_trees$field_plottree_id)
          )
      }
    ## rename columns and flag
      true_positive_trees = true_positive_trees %>% 
        dplyr::rename(
          uas_tree_height_m = tree_height_m
          , uas_tree_id = treeID
          , field_uas_distance_m = distance_m
        ) %>% 
        dplyr::mutate(
          field_uas_group = "true positive"
        )
  }else{ # if there are spatially matched trees
    true_positive_trees = dplyr::tibble(
      uas_tree_id = as.character(NA)
      , field_plottree_id = as.character(NA)
      , uas_tree_height_m = as.numeric(NA)
      , field_tree_height_m = as.numeric(NA)
      , height_diff_pct = as.numeric(NA)
      , field_uas_distance_m = as.numeric(NA)
      , field_uas_group = as.character(NA)
    )
  }
  remove(potential_tree_pairs_temp)
  gc()
```

True positive detections

```{r fld-dta-tp-data}
#check data
  true_positive_trees %>%
    dplyr::summarise(
      n_rows = n()
      , unique_uas_trees = n_distinct(uas_tree_id)
      , unique_field_trees = n_distinct(field_plottree_id)
      , unique_pairs = n_distinct(field_plottree_id, uas_tree_id)
    ) %>% 
    kableExtra::kbl(caption = "results of true positive stem-uas tree matching") %>% 
    kableExtra::kable_styling()
```

### Combine with Commission and Omission

To determine UAS detected tree commissions (i.e. UAS detected trees within the overstory plot for which there was no stem-mapped tree pair; $Co$) this analysis used the 2023-06 BHEF overstory field survey plot center and plot radius of 11.35 m. UAS detected trees within this radius with an [estimated DBH](#local_mod) over 5 in (12.69 cm) that did not have a matched stem-mapped tree pair were considered commissions ($Co$).

Omissions ($Om$) are stem-mapped trees without a UAS detected tree match.

```{r fld-dta-comb, results='hide'}
field_uas_comparison = dplyr::bind_rows(
  ## true positive
    true_positive_trees
  ## omission
    , field_overstory %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(
        plot_id, field_plottree_id, field_tree_height_m
      ) %>% 
      dplyr::anti_join(
        true_positive_trees
        , by = dplyr::join_by(field_plottree_id)
      ) %>% 
      dplyr::mutate(
        field_uas_group = "omission"
      )
  ## commission
    , field_overstory %>% 
      sf::st_drop_geometry() %>% 
      # get field plot center locations
      dplyr::distinct(plot_id, utm_x, utm_y) %>% 
      sf::st_as_sf(coords = c("utm_x", "utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)) %>% 
      # buffer plot center with plot area 11.35m radius + mean error in true positive trees
      # sf::st_buffer(11.35+mean(true_positive_trees$field_uas_distance_m, na.rm = T)) %>% 
      # buffer plot center with plot area 11.35m radius
      sf::st_buffer(11.35) %>% 
      # join with uas tree points
      sf::st_join(
        treetops_sf_with_dbh %>% 
          dplyr::filter(
            !treeID %in% true_positive_trees$uas_tree_id
            & dbh_cm >= min_tree_dbh_cm
            & tree_height_m >= min_tree_height_m
          ) %>% 
          dplyr::select(treeID) %>% 
          dplyr::rename(uas_tree_id=treeID)
        , join = st_intersects
        , left = F # performs inner join to only keep uas trees and plots with a match
      ) %>% 
      sf::st_drop_geometry() %>% 
      dplyr::mutate(
        field_uas_group = "commission"
      )
) %>% 
  # attach uas data
  dplyr::left_join(
    treetops_sf_with_dbh %>%
      sf::st_set_geometry("geometry") %>% 
      dplyr::mutate(
        uas_tree_utm_x = sf::st_coordinates(geometry)[,1] #lon
        , uas_tree_utm_y = sf::st_coordinates(geometry)[,2] #lat
      ) %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(treeID, tree_height_m, dbh_cm, uas_tree_utm_x, uas_tree_utm_y) %>% 
      dplyr::rename(
        uas_tree_id = treeID
        , uas_tree_height_m = tree_height_m
        , uas_dbh_cm = dbh_cm
      )
    , by = dplyr::join_by(uas_tree_id)
  ) %>% 
  # attach field data
  dplyr::left_join(
    field_overstory %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(
        field_plottree_id, field_tree_height_m, field_dbh_cm, plot_id
        , tree_utm_x, tree_utm_y
      ) %>% 
      dplyr::rename(
        field_tree_utm_x = tree_utm_x
        , field_tree_utm_y = tree_utm_y
      )
    , by = dplyr::join_by(field_plottree_id)
  ) %>% 
  # update data
  dplyr::mutate(
    uas_tree_height_m = uas_tree_height_m.y
    , field_tree_height_m = field_tree_height_m.y
    , plot_id = dplyr::coalesce(plot_id.x, plot_id.y)
    , field_uas_group = factor(
      field_uas_group
      , ordered = T
      , levels = c(
        "true positive"
        , "commission"
        , "omission"
      )
    ) %>% forcats::fct_rev()
    , dbh_diff_cm = uas_dbh_cm - field_dbh_cm
    , tree_height_diff_m = uas_tree_height_m - field_tree_height_m
    , dbh_diff_pct = dbh_diff_cm/field_dbh_cm
    , height_diff_pct = tree_height_diff_m/field_tree_height_m
    , abs_dbh_diff_pct = abs(dbh_diff_pct)
    , abs_height_diff_pct = abs(height_diff_pct)
    , plot_lab = stringr::str_pad(plot_id, width = 4, side = "left", pad = "0")
  ) %>% 
  dplyr::relocate(field_uas_group) %>% 
  dplyr::select(-c(tidyselect::ends_with(".x"), tidyselect::ends_with(".y"))) %>% 
  # convert to imperial units
  calc_imperial_units_fn()

## write
write.csv(field_uas_comparison, paste0(delivery_dir,"/field_uas_comparison_data.csv"), row.names = F)

## aggregate data to plot level
field_uas_comparison_plot =
  field_uas_comparison %>% 
  dplyr::count(plot_id, plot_lab, field_uas_group) %>% 
  dplyr::group_by(plot_id, plot_lab) %>% 
  dplyr::mutate(
    pct = n/sum(n)
    , tot = sum(n)
    , tp_pct=max(ifelse(field_uas_group=="true positive",pct,0)) 
  ) %>% 
  dplyr::ungroup() %>% 
  # attach f score
  dplyr::inner_join(
    field_uas_comparison %>% 
      dplyr::count(plot_id, field_uas_group) %>% 
      dplyr::mutate(field_uas_group = dplyr::case_when(
        field_uas_group == "true positive" ~ "tp"
        , field_uas_group == "commission" ~ "co"
        , field_uas_group == "omission" ~ "om"
      )) %>% 
      tidyr::pivot_wider(
        names_from = field_uas_group
        , values_from = n
        , values_fill = 0
      ) %>% 
      dplyr::mutate(
        plot_f_score = dplyr::coalesce(
          2 * ( (tp/(tp+om)) * (tp/(tp+co)) ) / ( (tp/(tp+om)) + (tp/(tp+co)) )
          , 0
        )
      ) %>% 
      dplyr::select(plot_id, plot_f_score)
    , by = dplyr::join_by(plot_id)
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Summary of Validation Data

### True Positive, Commission, Ommission

Summary of tree true positive ($TP$), commission ($Co$), and omission ($Om$) detection by BHEF overstory plot.

```{r fld-dta-sum-grp}
field_uas_comparison_plot %>% 
ggplot(
  mapping = aes(x = pct, y = reorder(plot_lab, tp_pct), fill=field_uas_group, group=field_uas_group)
) +
  geom_col(
    width = 0.7, alpha=0.8
  ) +
  geom_text(
    mapping = aes(
        label = scales::percent(ifelse(pct>=0.12,pct,NA), accuracy = 1)
        , fontface = "bold"
      )
    , position = position_stack(vjust = 0.5)
    , color = "black", size = 2.3
  ) +
  scale_fill_viridis_d(option = "cividis") +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(
    fill = ""
    , y = ""
    , x = "Percent of Trees"
    , title = "UAS and Stem-Mapped Tree Validation Summary"
    , subtitle = "by overstory plot"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , legend.title = element_text(size=7)
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.title.y = element_blank()
    , axis.text.x = element_blank()
    , axis.text.y = element_text(color = "black",size=10, face = "bold")
    , axis.ticks.x = element_blank()
  ) +
  guides(
    fill = guide_legend(reverse = T, override.aes = list(alpha = 0.9))
  )  
      
```

### F-score

```{r fld-dta-sum-f}
field_uas_comparison_plot %>% 
  dplyr::distinct(plot_lab, plot_f_score) %>% 
ggplot(
  mapping = aes(x = plot_f_score, y = reorder(plot_lab, plot_f_score), fill=plot_f_score, label = scales::comma(plot_f_score, accuracy = 0.01))
) +
  geom_vline(
    xintercept = field_uas_comparison_plot %>% dplyr::distinct(plot_lab, plot_f_score) %>% 
      dplyr::pull(plot_f_score) %>% median()
    , color = "gray44"
    , linetype = "dashed"
  ) +
  geom_col(
    width = 0.7, alpha=0.8
  ) +
  geom_text(
    color = "black", size = 3.5
    , hjust = -0.1
  ) +
  annotate(
    "text"
    , x = field_uas_comparison_plot %>% dplyr::distinct(plot_lab, plot_f_score) %>% 
      dplyr::pull(plot_f_score) %>% median()
    , y = 0 
    , label = paste0(
      "median: "
      , field_uas_comparison_plot %>% dplyr::distinct(plot_lab, plot_f_score) %>% 
      dplyr::pull(plot_f_score) %>% median() %>% scales::comma(accuracy = 0.01)
    )
    , hjust = -0.1
    , vjust = -1
  ) +
  scale_fill_viridis_c(option = "mako", direction = -1) +
  labs(
    fill = ""
    , y = ""
    , x = "F-Score"
    , title = "UAS and Stem-Mapped Tree F-Score Summary"
    , subtitle = "by overstory plot"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.title.y = element_blank()
    , axis.text.x = element_blank()
    , axis.text.y = element_text(color = "black",size=10, face = "bold")
    , axis.ticks.x = element_blank()
  )
    
```

### Height vs. DBH of $Tp$, $Co$, $Om$

```{r fld-dta-sum-htdbh}
field_uas_comparison %>% 
  dplyr::mutate(
    dbh_temp = dplyr::coalesce(field_dbh_in, uas_dbh_in)
    , ht_temp = dplyr::coalesce(field_tree_height_ft, uas_tree_height_ft)
  ) %>% 
  ggplot(
    mapping = aes(x = ht_temp, y = dbh_temp, color = field_uas_group)
  ) +
    geom_point(
      mapping = aes(shape = field_uas_group)
      , alpha=0.8
      , size=2
    ) +
    scale_color_viridis_d(option = "cividis") +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    scale_y_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      color = "detection"
      , shape = "detection"
      , y = "DBH (in)"
      , x = "Tree Ht. (ft)"
      , title = "UAS and Stem-Mapped Tree Validation Summary"
      , subtitle = "height and DBH relationship"
    ) +
    theme_light() +
    theme(
      legend.position = "top"
      , legend.direction  = "horizontal"
      , legend.title = element_blank()
    ) +
    guides(
      color = guide_legend(reverse = T, override.aes = list(alpha = 0.9, size = 5))
      , shape = guide_legend(reverse = T, override.aes = list(alpha = 0.9, size = 5))
    )  
```

### Height and DBH Distribution $Tp$, $Co$, $Om$

```{r fld-dta-sum-htdbh-dist}
field_uas_comparison %>% 
  dplyr::mutate(
    dbh = dplyr::coalesce(field_dbh_in, uas_dbh_in)
    , height = dplyr::coalesce(field_tree_height_ft, uas_tree_height_ft)
  ) %>% 
  dplyr::select(dbh, height, field_uas_group) %>% 
  tidyr::pivot_longer(cols = -c(field_uas_group), names_to = "metric", values_to = "value") %>% 
  dplyr::group_by(field_uas_group,metric) %>% 
  dplyr::mutate(
    metric = dplyr::case_when(
      metric == "dbh" ~ "DBH (in)"
      , metric == "height" ~ "Height (ft)"
    )
    , n_rows = dplyr::n()
    , plot_lab = paste0(
      field_uas_group
      ," (n=", scales::comma(n_rows,accuracy=1),")"
    )
  ) %>% 
  ggplot(mapping = aes(x = value, y = plot_lab, fill = field_uas_group)) +
    geom_violin(outlier.shape = NA) +
    geom_boxplot(width = 0.1, outlier.shape = NA, color = "gray66") +
    facet_grid(cols = vars(metric), scales = "free_x") +
    scale_fill_viridis_d(option = "cividis") +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      fill = ""
      , y = ""
      , x = ""
      , title = "UAS and Stem-Mapped Tree Validation Summary"
      , subtitle = "height and DBH distribution comparison"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
      , strip.text = element_text(color = "black", size = 12)
      , strip.background = element_rect(fill = "gray88")
    )
```

### Detected Overstory ($TP$) Height Difference

Detected overstory tree ($TP$) height reliability. Where overstory trees considered in this analysis had a tree height >= `r scales::comma(min_tree_height_m*3.28, accuracy = 0.1)` ft and a tree DBH >= `r scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)` in.

```{r fld-dta-sum-ht}
field_uas_comparison %>%
  dplyr::filter(field_uas_group == "true positive") %>%
  dplyr::group_by(plot_lab) %>% 
  dplyr::mutate(
    reorder_y = median(height_diff_pct, na.rm=T)
    , color_box = reorder_y<0 
    , n_rows = dplyr::n()
    , plot_lab = paste0(
      plot_lab
      ," (n=", scales::comma(n_rows,accuracy=1),")"
    )
  ) %>% 
  ggplot(mapping = aes(x = height_diff_pct, y = reorder(plot_lab, reorder_y) )) + # height_diff_pct
    geom_vline(xintercept = 0, color = "gray22") +
    geom_boxplot(mapping = aes(fill = color_box), alpha = 0.8, width = 0.7, outlier.shape = NA) +
    scale_fill_manual(values = c("steelblue", "coral")) +
    geom_point(color = "black", shape = ".") +
    scale_x_continuous(
      labels = scales::percent_format()
      , breaks = scales::extended_breaks(n=8)
      , limits = c(
        -max(field_uas_comparison$abs_height_diff_pct, na.rm = T)
        , max(field_uas_comparison$abs_height_diff_pct, na.rm = T)
      )
    ) +
    labs(
      fill = ""
      , y = ""
      , x = "Percent Difference in Height"
      , title = "Detected Overstory Height Difference"
      , subtitle = "by overstory plot"
      , caption = "-values = UAS<field | +values = UAS>field"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
    )
```

### Detected Overstory ($TP$) DBH Difference

Detected overstory tree ($TP$) DBH reliability. Where overstory trees considered in this analysis had a tree height >= `r scales::comma(min_tree_height_m*3.28, accuracy = 0.1)` ft and a tree DBH >= `r scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)` in.

```{r fld-dta-sum-dbh}
field_uas_comparison %>%
  dplyr::filter(field_uas_group == "true positive") %>%
  dplyr::group_by(plot_lab) %>% 
  dplyr::mutate(
    reorder_y = median(dbh_diff_pct, na.rm=T)
    , color_box = reorder_y<0 
    , n_rows = dplyr::n()
    , plot_lab = paste0(
      plot_lab
      ," (n=", scales::comma(n_rows,accuracy=1),")"
    )
  ) %>% 
  ggplot(mapping = aes(x = dbh_diff_pct, y = reorder(plot_lab, reorder_y) )) + # dbh_diff_pct
    geom_vline(xintercept = 0, color = "gray22") +
    geom_boxplot(mapping = aes(fill = color_box), alpha = 0.8, width = 0.7, outlier.shape = NA) +
    scale_fill_manual(values = c("steelblue", "coral")) +
    geom_point(color = "black", shape = ".") +
    scale_x_continuous(
      labels = scales::percent_format()
      , breaks = scales::extended_breaks(n=8)
      , limits = c(
        -max(field_uas_comparison$abs_dbh_diff_pct, na.rm = T)
        , max(field_uas_comparison$abs_dbh_diff_pct, na.rm = T)
      )
    ) +
    labs(
      fill = ""
      , y = ""
      , x = "Percent Difference in DBH"
      , title = "Detected Overstory DBH Difference"
      , subtitle = "by overstory plot"
      , caption = "-values = UAS<field | +values = UAS>field"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
    )

```

### Detected Overstory ($TP$) Reliability

Overstory trees considered in this analysis had a tree height >= `r scales::comma(min_tree_height_m*3.28, accuracy = 0.1)` ft and a tree DBH >= `r scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)` in.

```{r fld-dta-sum-rmse, results='asis'}
dbh_f_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(field_dbh_in)
dbh_u_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(uas_dbh_in)
ht_f_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(field_tree_height_ft)
ht_u_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(uas_tree_height_ft)

data.frame(
  mae_dbh = Metrics::mae(
    dbh_f_temp
    , dbh_u_temp
  )
  , mape_dbh = Metrics::mape(
    dbh_f_temp
    , dbh_u_temp
  )
  , rmse_dbh = Metrics::rmse(
    dbh_f_temp
    , dbh_u_temp
  )
    ## height
  , mae_height = Metrics::mae(
    ht_f_temp
    , ht_u_temp
  )
  , mape_height = Metrics::mape(
    ht_f_temp
    , ht_u_temp
  )
  , rmse_height = Metrics::rmse(
    ht_f_temp
    , ht_u_temp
  )
) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  tidyr::separate_wider_delim(cols = name, delim = "_", names = c("error", "metric")) %>% 
  tidyr::pivot_wider(names_from = error, values_from = value) %>% 
  dplyr::mutate(
    metric = dplyr::case_when(
      metric == "dbh" ~ "DBH (in)"
      , metric == "height" ~ "Height (ft)"
    )
    , n = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% nrow()
  ) %>% 
    kableExtra::kbl(
      caption = "Detected overstory tree height and DBH prediction performance"
      , col.names = c(
        " "
        , "Mean Abs. Error"
        , "Mean Abs. Percent Error"
        , "Root Mean Squared Error"
        , "N"
      )
      , digits = 2
    ) %>% 
    kableExtra::kable_styling()

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

### Detected Overstory ($TP$) Distribution Comparison

```{r fld-dta-sum-dist}
field_uas_comparison %>% 
  dplyr::filter(field_uas_group=="true positive") %>% 
  dplyr::select(c(
    tidyselect::ends_with("_dbh_in")
    , tidyselect::ends_with("_height_ft")
  )) %>% 
  dplyr::rename_with(
    .fn = function(x){dplyr::case_when(
      stringr::str_ends(x,"_dbh_in") ~ stringr::str_replace(x,"_dbh_in","_dbh")
      , stringr::str_ends(x,"_tree_height_ft") ~ stringr::str_replace(x,"_tree_height_ft","_height")
      , TRUE ~ x
    )}    
  ) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  tidyr::separate_wider_delim(
    cols = name
    , delim = "_"
    , names = c("source", "metric")
  ) %>% 
  dplyr::mutate(
    metric = dplyr::case_when(
      metric == "dbh" ~ "DBH (in)"
      , metric == "height" ~ "Height (ft)"
    )
    , source = dplyr::case_when(
      source == "uas" ~ "UAS"
      , source == "field" ~ "Stem Map"
    )
    , n = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% nrow()
  ) %>% 
  ggplot(mapping = aes(x = value, y = source )) +
    geom_boxplot(mapping = aes(fill = source), alpha = 0.8, width = 0.7, outlier.shape = NA) +
    # geom_point(color = "black", shape = ".") +
    facet_grid(cols = vars(metric), scales = "free_x") +
    scale_fill_manual(values = c("seagreen", "slateblue")) +
    labs(
      fill = ""
      , y = ""
      , x = ""
      , title = "Detected Overstory Distribution Comparison"
      , subtitle = paste0(
        "tree height >= "
        , scales::comma(min_tree_height_m*3.28, accuracy = 0.1)
        , " ft and tree DBH >= "
        , scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)
        , " in"
      )
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
      , strip.text = element_text(color = "black", size = 12)
      , strip.background = element_rect(fill = "gray88")
    )

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Commission and Omission Locations

```{r om-co-dta}
field_uas_comparison = field_uas_comparison %>% 
  dplyr::mutate(
    x = ifelse(!is.na(uas_tree_utm_x), uas_tree_utm_x, field_tree_utm_x)
    , y = ifelse(!is.na(uas_tree_utm_x), uas_tree_utm_y, field_tree_utm_y)
  ) %>% 
  sf::st_as_sf(
    coords = c("x", "y")
    , crs = sf::st_crs(treetops_sf_with_dbh)
    , remove=T
  )
```

View omissions and commissions on satellite imagery

```{r mapview-om-co, fig.height = 5}
field_overstory %>% 
  sf::st_drop_geometry() %>% 
  dplyr::distinct(plot_id, utm_x, utm_y) %>% 
  sf::st_as_sf(coords = c("utm_x", "utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)) %>% 
  sf::st_buffer(11.35) %>% 
  dplyr::mutate(x=1) %>% 
mapview::mapview(
  col.regions = "blue", layer.name = "plot", alpha.regions = 0.6
  , popup = F
) +
mapview::mapview(
  field_uas_comparison %>% dplyr::filter(field_uas_group %in% c("commission", "omission"))
  , zcol = "field_uas_group"
  , col.regions = viridis::cividis(n=3)[1:2]
  , cex = 2
  , alpha.regions = 0.8
  , layer.name = "validation"
) +
mapview::mapview(
  harvests
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
)

```

### Othomosaic Commission and Omission Locations

Define function to plot orthomosaic + CHM + tree points + plot

```{r plt-fn-plts}
######################################################################################
# function to plot ortho + chm + points + stand
######################################################################################
ortho_plt_fn = function(my_plotid){
  # plot vector
    plot_temp = field_overstory %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(plot_id == my_plotid) %>% 
      dplyr::group_by(plot_id) %>% 
      dplyr::summarise(
       utm_x = dplyr::first(utm_x, na_rm = T), utm_y = dplyr::first(utm_y, na_rm = T) 
      ) %>% 
      dplyr::mutate(
        plot_lab = stringr::str_pad(plot_id, width = 4, side = "left", pad = "0")
      ) %>% 
      sf::st_as_sf(coords = c("utm_x", "utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)) %>% 
      sf::st_buffer(11.35)
# convert to stars
  ortho_st = ortho_rast %>%  
    terra::subset(subset = c(1,2,3)) %>%
    terra::crop(
      # stand %>% 
      plot_temp %>%
        sf::st_buffer(5) %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        terra::vect()
    ) %>% 
    # terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
    stars::st_as_stars()
  
  # convert to rgb
  ortho_rgb <- stars::st_rgb(
    ortho_st[,,,1:3]
    , dimension = 3
    , use_alpha = FALSE
    # , stretch = "histogram"
    , probs = c(0.005, 0.995)
    , stretch = "percent"
  )
  # ggplot
  plt_rgb <- ggplot() +
    stars::geom_stars(data = ortho_rgb[]) +
    scale_fill_identity(na.value = "transparent") + # !!! don't take this out or RGB plot will kill your computer
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(
      x = ""
      , y = ""
    ) +
    theme_void()
  
  # return(plt_rgb)
  ## combine all plot elements
  plt_combine =
    plt_rgb +
    ggnewscale::new_scale_fill() +
    geom_tile(
      data = chm_rast %>% 
        terra::crop(
          plot_temp %>% 
          terra::vect()
        ) %>% 
        terra::mask(
          plot_temp %>% 
          terra::vect()
        ) %>% 
        `*`(3.28) %>% # transform to feet
        as.data.frame(xy=T) %>% 
        rename(f=3)
      , mapping = aes(x=x,y=y,fill=f)
    ) +
    scale_fill_viridis_c(
      option="plasma", alpha = 0.6, name = "canopy ht. (ft)"
      , limits = c(
        min(terra::values(chm_rast), na.rm = T)*3.28
        , max(terra::values(chm_rast), na.rm = T)*3.28
      )
    ) +
    geom_sf(
      data = plot_temp
      , alpha = 0
      , lwd = 1.2
      , color = "blue"
    ) +
    ggnewscale::new_scale_fill() +
    geom_sf(
      data = field_uas_comparison %>% 
        dplyr::filter(plot_id == my_plotid)
      , mapping = aes(fill = field_uas_group)
      , color = "black"
      , shape = 21
      , size = 2
    ) +
    # scale_fill_manual(values = viridis::cividis(n=3)[1:2], name = "trees") +
    scale_fill_viridis_d(option = "cividis", name = "trees", drop = F) +
    labs(
      title = paste0(
        "Plot ID: "
        , field_uas_comparison %>% 
          dplyr::filter(plot_id == my_plotid) %>% 
          dplyr::pull(plot_lab) %>% 
          unique()
      )
    ) +
    theme(
      legend.position = "top"
      , legend.direction = "horizontal"
      , legend.margin = margin(0,0,-1,0)
      , legend.text = element_text(size = 8)
      , legend.title = element_text(size = 8, face = "bold")
      , plot.title = ggtext::element_markdown(size = 10, hjust = 0.5)
    ) + 
    guides(
      fill = guide_legend(override.aes = list(size = c(5,NULL)))
    )  
  return(plt_combine)
}
# ortho_plt_fn(2)
```

Plot orthomosaic + CHM + tree points + plot for all plots with at least one omisson or commission

```{r plt-fn-plts-doit}
# map over plots with omissions and commisssions
field_uas_plts = field_uas_comparison %>% 
  dplyr::filter(field_uas_group %in% c("commission", "omission")) %>% 
  dplyr::pull(plot_id) %>% 
  unique() %>% 
  sort() %>% 
  purrr::map(ortho_plt_fn)
```

Combine plots

```{r plt-fn-plts-comb, fig.height=19}
patchwork::wrap_plots(
    field_uas_plts[]
    , ncol = 4
  ) +
  patchwork::plot_layout(guides="collect") &
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# check_ls_size_fn(ls()) %>% dplyr::slice_head(n=20)
remove(select_best_tree_pair_fn, true_positive_trees, field_uas_plts, ortho_plt_fn)
remove(list = ls()[grep("_temp",ls())])
gc()
```

